### 2020-02-06 text/20/panel-counts.Rmd
Added panel-counts

---
title: "Blog post: Gender distribution of panel discussions"
author: "Steve Simon"
source: "New"
date: "2020-02-06"
category: Blog post
tags: Human side of statistics, R software
output: html_document
---

I was curious, while I was attending the Joint Statistical Meetings (JSM) in August 2019 about the gender distribution of the panel discussions. There had been a lot of controversy at several scientific meetings about "manels," panel discussions that had all or almost all males. So I copied the JSM program into a text file and wrote a short R program to analyze this.

<!---More--->

Here are the R libraries that I need for this program. The suppressWarnings and suppressMessages functions help to keep the output clean.

```{r read}
suppressWarnings(suppressMessages(library(gender)))
suppressWarnings(suppressMessages(library(magrittr)))
suppressWarnings(suppressMessages(library(tidyverse)))
```

```{r data, echo=FALSE}
raw_text <- '
2019 Joint Statistical Meetings - Statistics: Making an Impact - Denver, Colorado
Statistics: Making an Impact
July 27  August 1, 2019
Colorado Convention Center
JSM 2019 Online Program
Return to main conference page
Online Program Home
 	My Program
    Quick Guide (PDF)	Register Now!
Legend:
CC = Colorado Convention Center   H = Hyatt Regency Denver at Colorado Convention Center 
* = applied session       ! = JSM meeting theme

Activity Details

15 !	Sun, 7/28/2019, 2:00 PM - 3:50 PM	CC-503
Artificial Intelligence for Data Science  Invited Panel
Council of Chapters
Organizer(s): Jason H Moore, University of Pennsylvania
Chair(s): Rebecca Hubbard, University of Pennsylvania
2:05 PM	Artificial Intelligence for Data Science 
Panelists:	Jason H Moore, University of Pennsylvania
Joel Dudley, Icahn School of Medicine at Mount Sinai
Larry Hunter, University of Colorodo Denver
3:40 PM	Floor Discussion

16 !	Sun, 7/28/2019, 2:00 PM - 3:50 PM	CC-205
To the Point: Critical Skills and Knowledge to Be Successful in Academia, Industry and the Government 
 Invited Panel
Committee on Career Development, Section on Statistical Consulting, Section on Statistics and Data 
Science Education
Organizer(s): Adin-Cristian Andrei, Northwestern University
Chair(s): Adin-Cristian Andrei, Northwestern University
2:05 PM	To the Point: Critical Skills and Knowledge to Be Successful in Academia, Industry and 
the Government 
Panelists:	David Morganstein, Westat
Jeri Metzger Mulrow, Bureau of Justice Statistics, Department of Justice, Washington, DC
Joan Chmiel, Northwestern University, Chicago, IL
John Bailer, Miami University
Mary J Kwasny, Northwestern University
3:45 PM	Floor Discussion

57 * !	Sun, 7/28/2019, 4:00 PM - 5:50 PM	CC-205
Women Transitioning into Leadership  Invited Panel
Caucus for Women in Statistics
Organizer(s): Kathy Panageas, Memorial Sloan Kettering Cancer Center
Chair(s): Kathy Panageas, Memorial Sloan Kettering Cancer Center
4:05 PM	Women Transitioning into Leadership 
Panelists:	Janet Wittes, Statistics Collaborative, Inc
Marcey Abate Hoover, Sandia National Laboratories
Melissa Begg, Columbia University 
Nancy Geller, National Institutes of Health, NHLBI
Stacy Lindborg, Biogen
5:45 PM	Floor Discussion

58 * !	Sun, 7/28/2019, 4:00 PM - 5:50 PM	CC-102
Artificial Intelligence (AI) in Healthcare and Medical Research in the Big Data Era  Invited Panel
Section on Statistics in Genomics and Genetics, WNAR, International Chinese Statistical Association
Organizer(s): Haiyan Huang, University of California, Berkeley
Chair(s): Ruixiao Lu, Genomic Health, Inc.
4:05 PM	Artificial Intelligence (AI) in Healthcare and Medical Research in the Big Data Era 
Panelists:	Bin Chen, Michigan State University
Haoda Fu, Eli Lilly and Company
Jean Yee Hwa Yang, University of Sydney, Australia
Jing Huang, Veracyte Inc.
Ying Lu, Stanford University
5:45 PM	Floor Discussion

223 * !	Mon, 7/29/2019, 2:00 PM - 3:50 PM	CC-703
The P-Value Controversy: Where Do We Go from Here?  Invited Panel
Biopharmaceutical Section, Biometrics Section, ENAR
Organizer(s): Pranab K Mitra, Merck
Chair(s): Pranab K Mitra, Merck
2:05 PM	The P-Value Controversy: Where Do We Go from Here? 
Panelists:	David Gal, University of Illinois at Chicago
Jeffrey Blume, Vanderbilt University
Lisa Strug
Michael Lavine, University of Massachusetts, Amherst
Philip B. Stark, UC Berkeley
Xihong Lin, Harvard
3:40 PM	Floor Discussion

224 * !	Mon, 7/29/2019, 2:00 PM - 3:50 PM	CC-Four Seasons 1
Sexual Harassment and Assault  Confronting the Threat to Our Statistical Community  Invited Panel
ASA Task Force on Sexual Harassment and Assault, Committee on Women in Statistics, Caucus for 
Women in Statistics
Organizer(s): Leslie McClure, Drexel University
Chair(s): Robert Santos, The Urban Institute
2:05 PM	Sexual Harassment and Assault: Confronting the Threat to Our Statistical Community 
Panelists:	Donna E LaLonde, ASA Committee on Women in Statistics
Emma Benn, Icahn School of Medicine at Mount Sinai
Leslie McClure, Drexel University
Sally C. Morton, Virginia Tech
3:45 PM	Floor Discussion

280 !	Tue, 7/30/2019, 8:30 AM - 10:20 AM	CC-503
Statistical Outreach and Awareness: How to Make an Impact  Invited Panel
Section for Statistical Programmers and Analysts
Organizer(s): Marianne Miller, Eli Lilly and Company
Chair(s): Adrian Coles, Eli Lilly and Co.
8:35 AM	Statistical Outreach and Awareness: How to Make an Impact 
Panelists:	Darius McDaniel, Emory
Jesse Chittams, University of Pennsylvania
Lillian Prince, Kent State University
Mark Ward, Purdue University
Renee Moore, Emory University
10:15 AM	Floor Discussion

281 * !	Tue, 7/30/2019, 8:30 AM - 10:20 AM	CC-205
When Statistical Methods Impact Policy  Invited Panel
Health Policy Statistics Section, Biometrics Section
Organizer(s): Sherri Rose, Harvard Medical School
Chair(s): Samrachana Adhikari, New York University
8:35 AM	When Statistical Methods Impact Policy 
Panelists:	Dionne Price, Food and Drug Administration
Michael Baiocchi, Stanford University
Miguel Marino, Oregon Health Sciences University
Sherri Rose, Harvard Medical School
10:15 AM	Floor Discussion

282 !	Tue, 7/30/2019, 8:30 AM - 10:20 AM	CC-102
Data Science Education at the School Level  Invited Panel
International Association for Statistical Education, National Council of Teachers of Mathematics, Section 
on Statistics and Data Science Education
Organizer(s): Tim Erickson, Epistemological Engineering
Chair(s): Tim Erickson, Epistemological Engineering
8:35 AM	Data Science Education at the School Level 
Panelists:	Andee Rubin, TERC
Anna Fergusson, University of Auckland
Michelle Wilkerson, University of California at Berkeley
Rob Gould, ASA
William Finzer, Concord Consortium
10:15 AM	Floor Discussion

283	Tue, 7/30/2019, 8:30 AM - 10:20 AM	CC-603
Statistical Sleuths, Data Thugs, or Methodological Terrorists? Recent Stories from the New Field of Error 
Detection in the Published Literature  Invited Panel
Section on Statistics and Data Science Education
Organizer(s): Regina Nuzzo, American Statistical Association; Kristin Sainani, Stanford University
Chair(s): Regina Nuzzo, American Statistical Association
8:35 AM	Statistical Sleuths, Data Thugs, or Methodological Terrorists? Recent Stories from the 
New Field of Error Detection in the Published Literature 
Panelists:	Andrew W. Brown, Indiana University School of Public Health-Bloomington
James Heathers, Northeastern University
Kristin Sainani, Stanford University
Nick Brown, University of Groningen
10:10 AM	Floor Discussion

324 *	Tue, 7/30/2019, 10:30 AM - 12:20 PM	CC-205
The Juggling Collaborative Statistician: Which Balls to Drop?  Invited Panel
Section on Statistical Consulting, Committee on Applied Statisticians, Biometrics Section, Korean 
International Statistical Society
Organizer(s): Julia L Sharp, Colorado State University
Chair(s): Mary J Kwasny, Northwestern University
10:35 AM	The Juggling Collaborative Statistician: Which Balls to Drop? 
Panelists:	Alexandra Hanlon, University of Pennsylvania
Emily Griffith, North Carolina State University
Ji-Hyun Lee, University of Florida
Julia L Sharp, Colorado State University
Jungwha "Julia" Lee, Northwestern University
Mimi Kim, Albert Einstein College of Medicine
12:15 PM	Floor Discussion

325 *	Tue, 7/30/2019, 10:30 AM - 12:20 PM	CC-102
Building Future Leaders: Perspectives on Training in Ethics, Professionalism, and Leadership  Invited 
Panel
Council of Chapters
Organizer(s): John D Keighley, University of Kansas Medical Center
Chair(s): David Morganstein, Westat
10:35 AM	Building Future Leaders: Perspectives on Training in Ethics, Professionalism, and 
Leadership 
Panelists:	Gina-Maria Pomann, Duke University
Jo Wick, University of Kansas Medical Center
Jonathan Gelfond, University of Texas Health San Antonio
Pandurang Kulkarni, Eli Lilly & Company
12:10 PM	Floor Discussion

388 * !	Tue, 7/30/2019, 2:00 PM - 3:50 PM	CC-603
Building Bridges for Data Science Education  Invited Panel
Section on Statistics and Data Science Education, Section on Statistical Computing, Section on Statistical 
Learning and Data Science
Organizer(s): Mine Cetinkaya-Rundel, Duke University
Chair(s): Beth Chance, Cal Poly - San Luis Obispo
2:05 PM	Building Bridges for Data Science Education 
Panelists:	Mine Cetinkaya-Rundel, Duke University
Michael Posner, Villanova University
Jeff Forbes, Duke University
Andrea Danyluk, Williams College
3:45 PM	Floor Discussion

389 !	Tue, 7/30/2019, 2:00 PM - 3:50 PM	CC-703
Official Statistics at the Crossroads: Data Quality and Access in an Era of Heightened Privacy Risk  
Invited Panel
Government Statistics Section, Committee on National Statistics, NAS, Committee on Professional Ethics
Organizer(s): Michael Hawes, U.S. Census Bureau
Chair(s): Michael Hawes, U.S. Census Bureau
2:05 PM	Official Statistics at the Crossroads: Data Quality and Access in an Era of Heightened 
Privacy Risk 
Panelists:	John M. Abowd, U.S. Census Bureau
Katy Rother, Committee on Oversight, U.S. House of Representatives
Michael Davern, NORC
Rochelle (Shelly) Wilkie Martinez, U.S. Office of Management and Budget
3:40 PM	Floor Discussion

491 *	Wed, 7/31/2019, 10:30 AM - 12:20 PM	CC-703
Database Lock to Data Safety Monitoring Board Meeting  More Than a Click of a Button  Invited 
Panel
Section for Statistical Programmers and Analysts, Biopharmaceutical Section, Section on Statistical 
Consulting
Organizer(s): Vipin Arora, Eli Lilly and Company
Chair(s): Vipin Arora, Eli Lilly and Company
10:35 AM	Database Lock to Data Safety Monitoring Board Meeting  More Than a Click of a 
Button? 
Panelists:	David Prince, Axio Research 
Kevin Buhr, University of Wisconsin
Lisa Weissfeld, Stats Collaborative
Natasa Rajicic, Cytel Inc
12:15 PM	Floor Discussion

492 * !	Wed, 7/31/2019, 10:30 AM - 12:20 PM	CC-205
Data Fabrication and Falsification: Protecting the Credibility and Impacts of Surveys  Invited Panel
Government Statistics Section, Survey Research Methods Section, ENAR
Organizer(s): Linda J Young, USDA National Agricultural Statistics Service
Chair(s): Kerrie Leslie
8:35 AM	Data Fabrication and Falsification: Protecting the Credibility and Impacts of Surveys 
Panelists:	James Dahlhamer, US Centers for Disease Control and Prevention
Jill DeMatteis , Westat
Linda J Young, USDA National Agricultural Statistics Service
12:15 PM	Floor Discussion

554 * !	Wed, 7/31/2019, 2:00 PM - 3:50 PM	CC-503
Interdisciplinary Research and Leadership: How to Make an Impact in the Data Science Age  Invited 
Panel
IMS, Section on Statistical Learning and Data Science, Royal Statistical Society
Organizer(s): Bin Yu, UC Berkeley
Chair(s): Bin Yu, UC Berkeley
2:05 PM	Interdisciplinary Research and Leadership: How to Make an Impact in the Data Science 
Age 
Panelists:	Alicia Carriquiry, Iowa State University
Christopher Genovese, Statistics, CMU
Hongyu Zhao, Yale
Jasjeet Sekhon, UC Berkeley
Simon Tavare, Inst of Cancer Dynamics and Statistics, Columbia University
Tamara Tamara Greasby, Data Science at The Trade Desk
3:45 PM	Floor Discussion

555 * !	Wed, 7/31/2019, 2:00 PM - 3:50 PM	CC-704
A Historical Perspective on the Application of Sampling Theory and Methods to Statistical  Invited 
Panel
Statistical Auditing Interest Group, History of Statistics Interest Group, Survey Research Methods Section
Organizer(s): John Hilton, Office of the Auditor General of Canada; Roger C. Pfaffenberger, Ryan, LLC
Chair(s): John Hilton, Office of the Auditor General of Canada
2:05 PM	A Historical Perspective on the Application of Sampling Theory and Methods to 
Statistical Auditing 
Panelists:	Donald Roberts, University of Illinois - Emeritus Professor Retired
Wendy Rotz, Grant Thornton
Richard Valliant, University of Maryland - Emeritus Professor Retired
Ron Bartyczak, Internal Revenue Service - Retired
Alan H. Kvanli, University of North Texas
3:45 PM	Floor Discussion

556 !	Wed, 7/31/2019, 2:00 PM - 3:50 PM	CC-205
Census 2020:  Invited Panel
Committee on Scientific Freedom and Human Rights
Organizer(s): Robin Mejia, Carnegie Mellon University
Chair(s): Robin Mejia, Carnegie Mellon University
2:05 PM	Census 2020 
Panelists:	Connie Citro, The National Academies of Sciences, Engineering, and Medicine
Edward Kissam, CIRS
Hansi Lo Wang, National Public Radio
3:45 PM	Floor Discussion

597 * !	Thu, 8/1/2019, 8:30 AM - 10:20 AM	CC-503
Vision 2020: Making Impact with Statistics in the Era of Data Science  Invited Panel
Committee of Presidents of Statistical Societies, ENAR, Section on Statistical Learning and Data Science
Organizer(s): Huixia Judy Wang, The George Washington University
Chair(s): Bhramar Mukherjee, University of Michigan
8:35 AM	Vision 2020: Making Impact with Statistics in the Era of Data Science 
Panelists:	Hadley Wickham, RStudio
Jeffrey Leek, Johns Hopkins Bloomberg School of Public Health
John Quackenbush, Harvard University
Rachel Schutt, BlackRock
Tian Zheng, Columbia University
Xiao-Li Meng, Harvard University
10:15 AM	Floor Discussion

639 * !	Thu, 8/1/2019, 10:30 AM - 12:20 PM	CC-503
Women in Data Science: a Small N Sample  Invited Panel
Section for Statistical Programmers and Analysts, Section on Statistical Learning and Data Science, 
Caucus for Women in Statistics
Organizer(s): Maria A Terres, Waymo
Chair(s): Maria A Terres, Waymo
10:35 AM	Women in Data Science: a Small N Sample 
Panelists:	Moorea Brega, Pattern Ag
Molly Davies, Stitch Fix
Mary Beth Broadbent, Google/YouTube
Cheryl Flynn, AT&T Research Labs
Clara Yuan, Convoy Inc.
12:05 PM	Floor Discussion

640 * !	Thu, 8/1/2019, 10:30 AM - 12:20 PM	CC-205
Providing Open Standards for Mobile Phone Data Statistics  Invited Panel
International Statistical Institute, Stats. Partnerships Among Academe Indust. & Govt. Committee, 
Statistics Without Borders
Organizer(s): May Offermans, Statistics Netherlands
Chair(s): Tracey Li, Flowminder
10:35 AM	Providing Open Standards for Mobile Phone Data Statistics 
Panelists:	May Offermans, Statistics Netherlands
Albrecht Wirthmann, EUROSTAT
Samantha Watson, Flowminder Foundation
12:15 PM	Floor Discussion

?
American Statistical Association
732 North Washington Street
Alexandria, VA 22314
(703) 684-1221
meetings@amstat.org

Copyright  American Statistical Association

Privacy Policy | Conduct Policy | Previous JSMs
'
raw_text %<>% str_split("\n") %>% unlist
```

The text file has a consistent format that makes it easy to identify things. The first panelist is on a line starting with the text "Panelists:" and the last panelist appears just before the line with the text "Floor Discussion".

```{r extract}
i <- grep("^Panelists:", raw_text)
j <- grep("Floor Discussion", raw_text)
print(i)
print(j)
for (k in 1:length(i)) {
  cat(paste(raw_text[i[k]:j[k]], collapse="\n"))
  cat("\n\n")
}
full_name <- NULL
panel <- NULL
for (k in 1:length(i)) {
  full_name <- c(full_name, raw_text[i[k]:(j[k]-1)])
  panel <- c(panel, rep(k, j[k]-i[k]))
}
full_name %>%
  sub("^Panelists:\t", "", .) %>%
  sub(" .*", "", .) -> first_name
```

The gender package takes a first name and compares it to a database of names from Social Security that calculates the proportion of males and of females in that database who have a specific first name. Names like "Mary" and "John" will belong almost exclusively to one gender, but names like "Pat" will be more evenly split.

```{r gender}
identify_gender <- gender(first_name)
identify_gender %>%
  distinct(name, .keep_all=TRUE) %>%
  full_join(data.frame(panel=panel, name=first_name, stringsAsFactors=FALSE)) -> panel_demographics
```

The gender package will only report results on names that are found in their database, and this would exclude some (but not all) "foreign" names. Any of these cases, plus any cases where the gender package rates the probability of male/female somewhere in the ambiguous middle (between 20% and 80%) will be coded as unknown/ambiguous.

Here are the names with unknown gender. You could probably resolve some of these easily.

```{r count-unknown}
unknown_gender <- is.na(panel_demographics$gender)
panel_demographics %>%
  filter(unknown_gender) %>%
  pull(name) %>%
  unique %>%
  sort %>%
  paste(collapse=", ") %>%
  print
```

Here are the ambiguous names.

```{r count-ambiguous}
ambiguous_gender <- panel_demographics$proportion_male > 0.20 & panel_demographics$proportion_male < 0.80
panel_demographics %>%
  filter(ambiguous_gender) %>%
  pull(name) %>%
  unique %>%
  sort %>%
  paste(collapse=", ") %>%
  print
```

Code both the unknown and ambiguous names into the same category.

```{r code-unknown-and-ambiguous}
panel_demographics$gender[unknown_gender] <- "unknown/ambiguous"
panel_demographics$gender[ambiguous_gender] <- "unknown/ambiguous"
table(panel_demographics$gender)
```

It is very easy to get a count of female and male names by panel. The proportion of females excludes any names that were not cdoed as unknown/ambiguous.

```{r summarize}
panel_demographics %<>% arrange(panel, gender)
n_panels <- max(panel_demographics$panel)
name_list <- rep("", n_panels)
for (i in 1:n_panels) {
  name_list[i] <- paste(panel_demographics$name[panel_demographics$panel==i], collapse=", ")
}
panel_demographics %>%
  mutate(f=as.numeric(gender=="female")) %>%
  mutate(m=as.numeric(gender=="male")) %>%
  mutate(u=as.numeric(gender=="unknown/ambiguous")) %>%
  group_by(panel) %>%
  summarize(
    f=sum(f),
    m=sum(m),
    u=sum(u)) %>%
  mutate(p_f=round(f/(f+m), 2)) %>%
  bind_cols(tibble(name_list)) %>%
  arrange(p_f) %>%
  data.frame -> panel_summary
panel_summary %>%
  summarize_at(c("f", "m", "u"), sum) %>%
  mutate(p_f=round(f/(f+m), 2)) -> panel_summary_total_row
panel_summary         
panel_summary_total_row
```