### 2020-03-16 text/20/map-projections.Rmd
Added geographic facts about Colorado

---
title: "Blog post: Different map projections"
author: "Steve Simon"
source: "New"
date: "2020-03-15"
category: Blog post
tags: Graphical display
output: html_document
---

I have been drawing some maps in R and was struggling to figure out how to display things with the appropriate map projections. Here is some code where I experimented with a variety of approaches.

<!---More--->

The earth is a three dimensional object and there is no way to display it on a two dimensional computer screen without some distortion. Your goal is to minimize that distortion.

First things first. There are a variety of map options, and they have a variety of storage methods.

One of the [tutorials](https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html) I viewed had mentioned [Natural Earth](http://www.naturalearthdata.com/), an open source repository. You can get easy access to this information with two R packages: [rnaturalearth](https://cran.r-project.org/package=rnaturalearth) and [rnaturalearthdata](https://cran.r-project.org/package=rnaturalearthdata).

What these give you is the ability to draw outlines of any and every country. You also get the outlines of states/provinces/districts in many countries. There's a lot more, but I want to just show something simple now to get you started.

The traditional format for storing map information is shapefiles. The shapefile format was defined the 1990s and is a rather intimidating collection of multiple files. You can find a nice summary on the [Wikipedia page on shapefiles](https://en.wikipedia.org/wiki/Shapefile) or at a Library of Congress's [page on archival formats](https://www.loc.gov/preservation/digital/formats/fdd/fdd000280.shtml).

A newer format, simple features, has greater flexibility and a simpler structure. It's still not simple, and I have been able to delve into the inner workings to get important details extracted.

There is a package in R, [sf](https://cran.r-project.org/package=sf), that allows you to easily manipulate geographic data stored in a simple features format. This package has a nice series of vignettes, starting [here](https://r-spatial.github.io/sf/articles/sf1.html).

Please note that the rnaturalearthhires package is needed below, but it cannot be downloaded directly (at least on my system). Instead you have to type in the command

```{}
install.packages("rnaturalearthhires", repos="http://packages.ropensci.org", type="source")
```

So to get everthing to work, even for a simple example, you need a whole bunch of packages. Here are the ones I use in this example.

```{r setup}
suppressMessages(suppressWarnings(library(lubridate)))
suppressMessages(suppressWarnings(library(mapproj)))
suppressMessages(suppressWarnings(library(rgeos)))
suppressMessages(suppressWarnings(library(rjson)))
suppressMessages(suppressWarnings(library(rnaturalearth)))
suppressMessages(suppressWarnings(library(rnaturalearthdata)))
suppressMessages(suppressWarnings(library(sf)))
suppressMessages(suppressWarnings(library(stringr)))
suppressMessages(suppressWarnings(library(tidyverse)))
```

I'm going to read in the United States file using the ne_states function. The country="United States of America" parameter excludes other countries that have states, provinces, or districts. I further want to remove Alaska and Hawaii from the file. So sorry to our 49th and 50th states, but the graphs are much more manageable for these examples when I focus just on the 48 states. The str function tells you a bit of information about how the data is organized within the states data frame.

```{r us-states}
ne_states(country="United States of America", returnclass = "sf") %>%
  filter(postal != "HI" & postal != "AK") -> states
str(states)
```

There is a lot of information in the list above and more than half of it is just giving different names to each state. The last column in this data frame is geometry and this is going to be the last column of any data stored in the simple features format. It ends up being a very long string with latitude and longitude coordinates for one or more polygons. You need multiple polygons for a disconnected state like Michigan, or a state with a lot of islands, like Florida.

Similarly, the ne_countries function gets information about the boundaries of every country in the world.

```{r world-countries}
countries <- ne_countries(scale = "medium", returnclass = "sf")
str(countries)
```

Notice that this data frame has a lot of statistical information (population, area, etc.) for each country. You don't need that information for the examples shown here, but it is nice to have.

Here is a picture of the earth as it might look like if viewed from the sun. The north pole is showing because it is summertime in the northern hemisphere. All the countries that are easily visible in this perspective are outlined in black and the state of Colorado in the United States is outlined in red.

```{r, echo=FALSE}
projection_string <- "+proj=ortho +lon_0=-80 +lat_0=23.5 +ellps=WGS84 +no_defs"
states %>%
  filter(postal=="CO") -> co
countries %>%
  ggplot() + 
  geom_sf(fill="white") + 
  geom_sf(data=co, color="red", fill="white") +
  coord_sf(crs=projection_string)
```

Colorado is a very interesting state with a simple boundary. The sourthern edge is 37 degrees latitude. The northern edge is 41 degrees north latitude. The eastern edge is 102 degrees, 02 minutes, 48 seconds west longitude. The western edge is 109 degrees, 02 minutes, 48 seconds west longitude.

You can calculate the length of each of these boundaries with a bit of simple math. The radius of the earth is

$r=6,378$ meters

The circumference is

$2\pi r=40,074$ meters

Each degree of latitude would be

$40,074/360=111.3$ meters.

So the four degrees that you would have to travel to get from the southern border (37 degrees) to the northern border (41 degrees) is

$4*111.3=445.2$ meters

The difference between the northern and southern borders is 4 degrees, so the length of either the eastern or the western borders is

$(4/360)*40,074=445$ meters

The distance from the east to the west border is tricky. At the equator a degree of distance is 111.3 meters, but as you move away from the equator, the lines of latitude converge so a degree of longitude shrinks.