### 2020-04-17 text/20/codes-for-covid-19.Rmd
Added emcee-package.md

---
title: "Codes for COVID-19 in the electronic health record"
author: "Steve Simon"
source: "New"
date: "2020-04-15"
category: Blog post
tags: Electronic health records
output: html_document
---

I might be helping with a research project that involves selection of medical records associated with COVID-19. Here are some of the resources and suggestions that I received from a colleague at KUMC plus additional ones I found from a web search. I include links to any resources that I drew from.

I would greatly appreciate anyone who could offer additional information, especially about additional CPT codes.

<!---More--->

There are both ICD-10, CPT, and LOINC codes that might help in identifying COVID-19 patients in the electronic health record.

#### ICD-10

The ICD-10 code for COVID-19 is U07.1 (COVID-19, virus identified) or U07.2 (COVID-19, virus not identified). This code is in a range of values, U00-U49, that have been set aside for  new diseases of uncertain etiology or emergency use. Previous codes in this range had included SARS (U04) and vaping related disorder (U07.0).

THe two codes are a recent addition, and it looks like earlier cases used an ICD-10 code of B97.29 (See the AAFP article or compare the February 20, 2020 CDC guidelines to the April 1, 2020 guidelines).

There are several respiratory illnesses associated with COVID-19. Pneumonia is coded as J12.89, acute bronchitis as J20.8, non-acute bronchitis as J40, lower respiratory infection as J22, acute respiratory distress syndrome as J80, and other respiratory disorders as J98.8.

People being tested for COVID-19 could have a code of Z03.818 (Encounter for observation for suspected exposure to other biological agents ruled out), Z20.828 (Contact with and (suspected) exposure to other viral communicable diseases) or Z11.59 (Encounter for screening for other viral diseases). Some of these, of course, might end up being negative tests.

A code that defintely indicates that COVID-19 has been ruled out is Z03.818 (Encounter for observation for suspected exposure to other biological agents ruled out).

Signs and symptoms are often vague, but three that are worth looking at are R05 (Cough), R06.02 (Shortness of breath), and R50.9 (Fever, unspecified).

#### CPT

There is a CPT code for COVID-19 testing, 87635, which was implemented on March 13, 2020. Additional codes were added/modified on April 10, 2020. These are 86318, 86328, 86602, 86635, and 86769. The details of these tests is too complex to summarize here.

There are certain additional CPT codes for procedures that may indicate treatment for COVID-19. These would include codes for ventilators and radiology, for example. I do not have a good list of what these codes might be.

#### Sensitive versus precise searches

When you are trying to identify COVID-19 cases, you have to decide whether you want a sensitive search, one that is unlikely to miss any cases, but which might include more than a few false positives. The alternative is a precise search, which provides relatively strong assurance that anyone you identify is indeed a COVID-19 case, but which also might miss more than a few cases. The answer will depend on the purpose of your research. It is also worth noting the additional complications caused by the current shortage of tests in the United States combined with the large number of people who seem to have the infection but who are asymptomatic.

Different groups use different criteria for deciding who qualifies to be counted as a COVID-19 patient. This usually involves

1. a positive COVID-19 test, or 
2. close contact with other(s) with COVID-19 combined with certain clinical signs or symptoms, or 
3. a greater combination of these clinical signs or symptoms even without a positive test or known close contact.

#### LOINC

There is a long list of LOINC codes associated with COVID-19:

+ 94306-8
+ 94307-6
+ 94308-4
+ 94309-2
+ 94310-0
+ 94311-8
+ 94312-6
+ 94313-4
+ 94314-2
+ 94315-9
+ 94316-7
+ 94500-6
+ 94502-2
+ 94503-0
+ 94504-8
+ 94505-5
+ 94506-3
+ 94507-1
+ 94508-9
+ 94509-7
+ 94510-5
+ 94511-3
+ 94531-1
+ 94532-9
+ 94533-7
+ 94534-5
+ 94547-7
+ 94558-4
+ 94559-2
+ 94562-6
+ 94563-4
+ 94564-2
+ 94565-9
+ 94639-2
+ 94640-0
+ 94641-8
+ 94642-6
+ 94643-4
+ 94644-2
+ 94645-9
+ 94646-7
+ 94647-5
+ 94660-8
+ 94661-6

This list is probably incomplete as of the writing of this webpage, and it will almost certainly become more incomplete over time.

Many of these tests, I suspect, are new and not helpful for detecting early cases of COVID-19. Also, these are only for the test orders and not the test results.

#### Miscellaneous

There are multiple terms relating to COVID-19. The term COVID-19 itself refers to the disease. The virus that causes COVID-19 is called SARS-CoV-2 or SARS Coronavirus 2. It was earlier referred to as 2019 Novel coronavirus or 2019-nCoV.

#### Resources

American Academy of Family Physicians. Coding for the new coronavirus and COVID-19 illness, March 5, 2020, updated March 19, 2020. Available in [html format](https://www.aafp.org/journals/fpm/blogs/gettingpaid/entry/coronavirus_diagnosis_coding.html).

American Academy of Pediatrics. How to use ICD-10-CM, new lab testing codes for COVID-19, March 12, 2020. Avialable in [html format](https://www.aappublications.org/news/2020/03/12/coding031220).

American Medical Association. Special coding advice during COVID-19 public health emergency. Available in [pdf format](https://www.ama-assn.org/system/files/2020-04/covid-19-coding-advice.pdf).

American Medical Association. CPT Category I Pathology and Laboratory Codes for Severe
Acute Respiratory Syndrome Coronavirus 2. Available in [pdf format](https://www.ama-assn.org/system/files/2020-04/coronavirus-long-descriptors.pdf).

American Thoracic Society. Coding & Billing Quarterly, April 2020. Available in [pdf format](https://www.thoracic.org/about/newsroom/newsletters/coding-and-billing/resources/2020/cbq-april.pdf).

Centers for Disease Control and Prevention. CD-10-CM Official Coding Guidelines - Supplement. Coding encounters related to COVID-19 Coronavirus Outbreak. Effective: February 20, 2020. Available in [pdf format](https://www.cdc.gov/nchs/data/icd/ICD-10-CM-Official-Coding-Gudance-Interim-Advice-coronavirus-feb-20-2020.pdf).

Centers for Disease Control and Prevention. ICD-10-CM Official Coding and Reporting Guidelines April 1, 2020 through September 30, 2020. Available in [pdf format](https://www.cdc.gov/nchs/data/icd/COVID-19-guidelines-final.pdf).

Euorpean Center for Disease Prevention and Control. Case definition and European surveillance for COVID-19, as of 2 March 2020. Available in [html format](https://www.ecdc.europa.eu/en/case-definition-and-european-surveillance-human-infection-novel-coronavirus-2019-ncov).

LOINC. SARS Coronavirus 2. Available in [html format](https://loinc.org/sars-coronavirus-2/).