### 2020-04-24 text/20/ihme-covid-model.Rmd
Continued work on ihme-covid-model.

---
title: "Criticisms of the IHME COVID-19 model"
author: "Steve Simon"
source: "New"
date: "2020-04-23"
category: Blog post
tags: Epidemiology
output: html_document
---

I was asked to review the details of a prominent prediction model for the COVID-19 crisis. This model was developed by the Institute for Health Metrics and Evaluation and is commonly referred to as the IHME model. Here is a bit of background on the IHME model and some resources that discuss this model.

<!---More--->

#### Overview of the IHME model

To avoid too much confusion, I will use the term "IHME model" to refer to the COVID-19 prediction model developed by the Institute for Health Metrics and Evaluation and reserve the term "the Institute" to refer to the organization itself.

The Institute has been in business since 2007, so they are no Johnny-Come-Lately operation. The page that lists their staff members has well over 400 pictures. Not all of these people are working on COVID-19, I'm sure, but it appears to be a solid team effort.

There is a wonderful sense of openness and transparency about research these days, and the Institute is a good example of this. The Institute shares the code and (when it can) the data on their [github site](https://github.com/ihmeuw). This allows other researchers to borrow their work and easily adapt it as needed. When the Institute makes changes to their work, anyone can quickly download these updates.

#### IMHE documentation of their model

Murray CJL. Forecasting the impact of the first wave of the COVID-19 pandemic on hospital demand and deaths for the USA and European Economic Area countries. Preprint, 2020-04-21. Available in [pdf format](http://www.healthdata.org/sites/default/files/files/Projects/COVID/RA_COVID-forecasting-USA-EEA_042120.pdf).

+ At the time of the writing of this webpage, this was the most current description of the IHME model.

Murray CJL. Forecasting COVID-19 impact on hospital bed-days, ICU-days, ventilator-days and deaths by US state in the next 4 months. Preprint, 2020-03-26. Available in [pdf format](https://www.medrxiv.org/content/10.1101/2020.03.27.20043752v1.full.pdf).

+ This article describes an earlier model.

Xu B, Gutierrez B, Mekaru S, Sewalk K, Goodwin L, Loskill A, Cohn EL, Hswen Y, Hill SC, Cobo MM, Zarebski AE, Li S, Wu C-H, Hulland E, Morgan JD, Wang L, O'Brien K, Scarpino SV, Brownstein JS, Pybus OG, Pigott DM, Kraemer MUG. Epidemiological data from the COVID-19 outbreak, real-time case information. Scientific Data, 2020-03-24. doi:10.1038/s41597-020-0448-0. Available in [html format](https://www.nature.com/articles/s41597-020-0448-0) or [pdf format](https://www.nature.com/articles/s41597-020-0448-0.pdf).

+ This article describes how the initial datasets used by the IHME model were collected. In a rapidly changing environment, some of the details may have changed, and at least one of the links in the article points to a gitrhub page that no longer exists. But most of the links are good. The breadth of the data collection is very impressive and the authors do a good job documenting everything.

IHME COVID-19 model FAQs. Available in [html format](http://www.healthdata.org/covid/faqs).

+ This webpage talks in non-technical terms about some of the choices that the researchers chose for the IHME model.

#### Articles in the popular press

Aizenman N. 5 Key Facts Not Explained In White House COVID-19 Projections. National Public Radio 2020-04-01. Available in [html format](https://www.npr.org/sections/health-shots/2020/04/01/824744490/5-key-facts-the-white-house-isnt-saying-about-their-covid-19-projections).

+ The focus of this article is on the White House press briefing and the uncertain source of some of the numbers used in that briefing. There is, however, a good amount of detail about the IHME model.

This article contrasts a model based on infectious disease theory, the Imperial College model, to the IHME model and then goes on to explain some of the features of the IHME model.

#### More technical articles

Druke N, Silver N. How One Modeler Is Trying To Forecast The Toll Of COVID-19. FiveThirtyEight 2020-04-22. Available in [video format](https://fivethirtyeight.com/features/politics-podcast-how-one-modeler-is-trying-to-forecast-the-toll-of-covid-19/) and as a [podcast](https://podcasts.apple.com/us/podcast/model-talk-forecasting-the-toll-of-covid-19/id1077418457).

This is an interview with Dr. Chris Murray, the person in charge of the IHME model.

Begley S. Influential Covid-19 model uses flawed methods and shouldn’t guide U.S. policies, critics say. Stat 2020-04-17. Available in [html format](https://www.statnews.com/2020/04/17/influential-covid-19-model-uses-flawed-methods-shouldnt-guide-policies-critics-say/).

#### Articles in the peer-reviewed literature

Jewell NP, Lewnard JA, Jewell BL. Caution Warranted: Using the Institute for Health Metrics and Evaluation Model for Predicting the Course of the COVID-19 Pandemic. Ann Intern Med. 2020. Available in [html format](https://doi.org/10.7326/M20-1565).