### 2020-08-21 text/20/collaboration-videos.Rmd
Added collaboration-videos

---
title: "Setting the Stage: Statistical Collaboration Training Videos"
author: "Steve Simon"
source: "New"
date: 2020-08-20
categories:
- Recommended
tags:
- Human side of statistics
output: html_document
---

Julia Sharp, Emily Griffith, and Megan Higgs, Ann Hess. Setting the Stage: Statistical Collaboration Training Videos. Available in [html format](https://sites.google.com/view/juliasharp/other-resources/statistical-collaboration-videos).

<!---More--->

+ A series of ten short videos (the longest is 5 minutes, 41 seconds) with a simulated consulting experience shown. These videos offer some thoughts on how to handle difficult issues that might arise during consulting.