### 2020-11-06 text/20/markdown-papers.Rmd
Started work on rmarkdown-papers

---
title: "How I use R Markdown to write research papers"
source: "New"
categories:
- Blog post
tags:
- Writing research papers
output: html_document
---

I joined a writing group at UMKC and am making very good progress on a couple of research papers. The group meets once a week to talk about everyone's progress and to share ideas. One interesting suggestion was the use of [Notion](https://www.notion.so/), software that organizes all your myriad writing tasks (actually tasks in general). I shared how I use R Markdown for writing and promised that I'd provide some resources about it before the next meeting.

<!--more-->

R Markdown is an implementation of [Markdown](https://commonmark.org/help/) inside the R programming language. R Markdown also leverages features in [LaTex](https://www.latex-project.org/), [MathJax](https://www.mathjax.org/), and [Pandoc](https://pandoc.org/).

If you use the [RStudio integrated development environment](https://rstudio.com/) to run R, then you already have many of the features of R built-in. There are several R packages that use or extend R Markdown:

+ [bookdown](https://bookdown.org/)
+ [blogdown](https://alison.rbind.io/post/2017-06-12-up-and-running-with-blogdown/)
+ [knitr](https://yihui.org/knitr/)
+ [rmarkdown](https://cran.r-project.org/web/packages/rmarkdown/index.html)

I use R Markdown to produce early drafts of research publications. It simplifies writing in several different ways.

**Figures** R can produce beautiful graphs, and with R Markdown, you can insert these graphs anywhere in your document. You can't control page breaks easily in R Markdown, however. Here's an example of a simple graph.

```{r, echo=FALSE}
suppressMessages(suppressWarnings(library(tidyverse)))
x=seq(-1, 1, length=10000)
y <- sin(1/x)
data.frame(x, y) %>%
  ggplot(aes(x, y)) + geom_line()
```