### 2021-08-30 text/21/poisson-inferences.Rmd
Updated poisson-inferences.md

---
title: "Working with the Poisson distribution"
author: "Steve Simon"
source: "New"
date: "2020-08-30"
categories:
- Blog post
tags:
- Probability distributions
output: html_document
---

Many statistics, such as mortality statistics, are based on fundamental counting processes. If you want to understand these processes, you should start by trying to understand the Poisson distribution. Here's a brief introduction.

```{r setup}
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

<!---More--->

### Definition of a Poisson distribution

A random variable X, defined on the set of non-negative integers (0, 1, 2, ...), is said to have a Poisson distribution (X ~ Pois($\lambda$) if

$P[X=k]=\frac{\lambda^ke^{-k}}{k!}$ where $\lambda > 0$

### Mean and variance of a Poisson random variable.

Without too much difficulty, you can show that

$E[X]=\lambda$

$V[X]=\lambda$

### sums of Poisson random variables

If

$X_1$ ~ Pois($\lambda_1$) and

$X_2$ ~ Pois($\lambda_2$) and

$X_1$ and $X_2$ are independent, then

$X_1+X_2$ ~ Pois($\lambda_1+\lambda_2$).

### Visualization of Poisson probabilities

```{r bar-charts}
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 0.5)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=0.5")
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 2)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=2")
data.frame(x=0:25) %>%
  mutate(p=dpois(x, 8)) %>%
  ggplot(aes(x,p)) +
    geom_col() +
    expand_limits(y=1) +
    ggtitle("Poisson probabilities with lambda=8")
```

### Relationship to the Binomial distribution

If Y has a binomial distribution with n trials and probability of success p (X ~ Bin(n,p)), then for large n and small p, you can approximate Y with X ~ Pois($\lambda$=np).

### Inference for unknown value of lambda

If you collect data on a single Poisson random variable X where the parameter $\lambda$ is unknown, then you can estimate $lambda$ with a reasonable amout of precision with just that one value.

$\hat{\lambda}=X$

and a crude 95% confidence interval for $\lambda$ is

$X \pm 2 \sqrt{X}$

This confidence interval does poorly for small values of $\lambda$ because of the skewness of the Poisson distribution. it also can sometimes produce confidence intervals with a negative lower limit.

### Estimation of a rate

A count X divided by a measure of time, T, is called a rate (R=X/T). If you assume that

x ~ Pois($\lambda$T)

where T is known, but $\lambda$ is unknown, then a 95% confidence interval for $\lambda$ is

$\frac{X}{T} \pm 2 \frac{\sqrt{X}}{T}$

### An improved confidence interval

The natural log transformation helps produce a better confidence interval.

The general rule is that a nonlinear transformation leads to a similar change in the expected value, but a change in the variance that is related to the Taylor series expansion.

If X has an expected value, $\mu$ and a variance, $\sigma^2_x$, then g(X) can be written as

$g(X) \approx g(\mu)+(X-\mu)g'(\mu)$

which can be used to show that

$Var(g(X))=(g'(\mu))^2 \sigma^2_x$

If

X ~ Pois($\lambda T$) and

$Y = g(X) = ln(X)$, then

$g'(X)=\frac{1}{X}$, and

V[Y] $\approx (g'(\mu))^2 \sigma^2_x = (\frac{1}{\lambda T})^2 \lambda T = \frac{1}{\lambda T}$

An approximate 95% confidence interval for ln($\lambda  T$) is

$ln(X) \pm 2 \sqrt{\frac{1}{X}}$

Exponentiate both sides to get a confidence interval for $\lambda T$.

Divide by T to get a confidence interval for $\lambda$.

### Examples

```{r echo=FALSE}
x <- 17
n <- 1989
lo <- x-2*sqrt(x)
hi <- x+2*sqrt(x)
log_lo <- log(x)-2*sqrt(1/x)
log_hi <- log(x)+2*sqrt(1/x)
exp_lo <- exp(log_lo)
exp_hi <- exp(log_hi)
```

The Pennsylvania Department of Health has a web page oulining calculations for rates with two simple examples. The infant mortality rate in Butler County in 1984 was 17 deaths out of 1,989 births. This is actually a proportion, but let's treat it as a rate here.

The death rate is 17/ 1989 =`r 17/1989`. To avoid dealing with small numbers, let's quote this as a rate per 1,000 live births. 1000 * 17/ 1989 = `r 1000*17/1989`. if we wanted a confidence interval for $\lambda T$, it would be

17 $\pm$  2 $\sqrt{17}$ = `r lo` to `r hi`

Divide everything by 1,989 and multiply by 1,000 to get

`r lo/1.989` to `r hi/1.989` which you should round to get 
`r round(lo/1.989, 1)` to `r round(hi/1.989, 1)`

Let's rework this using the better formula.

The confidence limits on the log scale are

$ln(17) \pm 2 \sqrt{\frac{1}{17}}$ = `r log_lo` to `r log_hi`

Exponentiate these values to get

`r exp_lo` to `r exp_hi`

Divide by 1,989 and multiply by 1,000 to get

`r exp_lo/1.989` to `r exp_hi/1.989`

and don't forget to round this: `r round(exp_lo/1.989, 1)` to `r round(exp_hi/1.989, 1)`.

### References

http://www.pmean.com/07/ConfidenceIntervalForRate.html

https://www.health.pa.gov/topics/HealthStatistics/Statistical-Resources/UnderstandingHealthStats/Documents/Confidence_Intervals_for_a_Crude_Rate.pdf

https://en.wikipedia.org/wiki/Taylor_expansions_for_the_moments_of_functions_of_random_variables

https://en.wikipedia.org/wiki/Delta_method

https://en.wikipedia.org/wiki/Poisson_distribution

https://en.wikipedia.org/wiki/Binomial_distribution#Poisson_approximation

http://www.pmean.com/definitions/or.htm

http://www.pmean.com/01/oddsratio.html

http://www.pmean.com/04/Rates.html