### 2021-09-30 text/Rmd/glm-deviance/index.Rmd

---
title: "Deviance in a generalized linear model"
author: "Steve Simon"
source: "New"
date: "2020-09-20"
categories:
- Blog post
tags:
- Generalized linear model
output: html_document
---

Deviance is a measure computed for generalized linear models that can help you decide between two competing models. It is somewhat analagous to residual sums of squares in linear regression.

```{r setup}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

<!---More--->

### References

Eduardo Garcia Portugues. Section 5.5. Deviance, in Notes for Predictive Modeling, 2021-08-18. Available in [html format][port1]

Freddy Hernandez. Deviance in glm. RPubs, 2020-07-07. Available in [html format][hern1]

German Rodriguez. Generalized Linear Model Theory. Lecture Notes on Generalized Linear Models. Available in [pdf format][rodr1]

Kenneth Tay. What is deviance? Statistical Odds and Ends blog, 2019-03-27. Available in [html format][tay1]

[hern1]: https://rpubs.com/fhernanb/deviance_glm
[port1]: https://bookdown.org/egarpor/PM-UC3M/glm-deviance.html
[rodr1]: https://data.princeton.edu/wws509/notes/a2.pdf
[tay1]: https://statisticaloddsandends.wordpress.com/2019/03/27/what-is-deviance/