### 2021-10-02 text/Rmd/8-tips/index.Rmd
Fixed 1999 dates, part 1

---
title: "8 tips for doing data visualization right."
author: "Steve Simon"
source: "http://blog.pmean.com/8-tips/"
date: "2014-05-03"
category: Recommendation
tags: Graphical display
output: html_document
---

This slide show includes some examples of really bad (and a few really
good) graphics with some explanations of general principles for data
display.

<!---More--->

AllAnalytics -- Beth Schultz -- 8 Tips for Doing Data Visualization
Right. Available at:
<http://www.allanalytics.com/author.asp?_mc=sem%5Fotb%5Fedt%5Fallanppcm&section_id=1411&doc_id=261787&image_number=>1.

```{r}
library(knitr)
# include_graphics("../../web/images/8-tips01.png")
```