### 2021-10-17 text/04/early-stopping-animal.Rmd
Renamed files in 2004 folder

---
title: Early stopping in an animal study
author: Steve Simon
date: 2004-07-01
categories:
- Blog post
tags:
- Early stopping
output: html_document
---
I was helping someone with power calculations for an animal experiment.
There were several ways to estimate power and they all seemed to point
to the use of either 8 animals per group or 12 animals per group. I
explained that if she was more risk averse, she should use 12 animals
and if she were more work averse, she should use 8 animals. Her answer
surprised me a bit. She said that she liked the idea of using 12 animals
per group. That way, if the results achieved significance after only 8
animals, she could go back to the Animal Care and Use Committee and tell
that she saved a few animals.

This is an example of early stopping of a clinical trial (also described
as an interim analysis). Generally, if you think you might want to stop
a trial early, you should specify the conditions in your protocol prior
to the start of data collection. Without such a rule, any examination of
the data prior to the end of the study is a protocol deviation. As such,
it should be reported to the IRB (if it is human research) or the ACUC
(if it is animal research).

There are reasonable examples of why you might want to stop a study
early even without planning ahead for such an event. For example, if a
serious unexpected side effect appears, you might not want to offer the
new drug to any more patients.

But stopping the study early (like any other protocol deviation) does
reduce the credibility of the research findings, sometimes by a little
bit, sometimes by a lot.

Here's a little simulation that you can run in S-plus or R.

```{}
par(mar=c(4,5,0,0)+0.1,cex=2,pch="+")
f.p <- function(mua) {
  x1 <- rnorm(12)
  x2 <- mua+rnorm(12)
  p.value <- rep(NA,12)
  for (i in 2:12) {
    p.value[i] <- t.test(x1[1:i],x2[1:i])p.value
 }
 plot(p.value,type="b",ylim=0:1,cex=2)
 abline(h=0.05)
}
for (i in 1:100) {
  f.p(mua=1)
}
```

You'll see a variety of different patterns for the p-values.

![](http://www.pmean.com/images/04/EarlyStopping01.gif)

In the graph shown above the test never quite reaches statistical
significance.

![](http://www.pmean.com/images/04/EarlyStopping02.gif)

Here's one where it just reaches statistical significance at the last
sample.

![](http://www.pmean.com/images/04/EarlyStopping03.gif)

Here's one where it reaches statistical significance rather early. You
would think that it would be nice to stop the experiment early here, but
there is no guarantee that just because the p-value is below the
significant level early on, that it will stay below that level the rest
of the way.

![](http://www.pmean.com/images/04/EarlyStopping04.gif)

Notice that the above graph shows statistical significance early, but
then that significance disappears by the end of the experiment. Now this
doesn't happen very often. Out of 100 graphs only 4 or 5 showed this
type of pattern. Most of the time if the p-value crosses below the 0.05
threshold early, it stays below that threshold. But these exceptions
serve as a warning. If you stop the study early and make no adjustments
to your p-value, your results may be invalid.

The adjustments to the p-value (and confidence interval) are rather
complex. Here's some more S-plus (R) code that shows some of the
complexities of the situation. This program simulates an experiment with
12 animals per group, but looks at the p-value with after 4, 8, and 12
animals have been tested.

```{r, echo=TRUE}
n <- 5000
p4 <- rep(NA,n)
p8 <- rep(NA,n)
p12 <- rep(NA,n)
for (i in 1:n) {
  x1 <- rnorm(12)
  x2 <- rnorm(12)
  p4[i] <- t.test(x1[1:4],x2[1:4])$p.value
  p8[i] <- t.test(x1[1:8],x2[1:8])$p.value
  p12[i] <- t.test(x1[1:12],x2[1:12])$p.value
}
sum(p4<0.05)/n
sum(p8<0.05)/n
sum(p12<0.05)/n
sum(pmin(p4,p8,p12)<0.05)/n
```

Notice that each p-value considered in isolation maintains an alpha
level of approximately 0.05. But when you combine the p-values and allow
for statistical significance when either the n=4, n=8, or n=12 p-value
achieves significance, the alpha level more than doubles.

**Further reading:**

Ludbrook J. Interim analyses of data as they accumulate in laboratory
experimentation. BMC Med Res Methodol 2003: 3(1); 15.
[Medline](http://www.ncbi.nlm.nih.gov/entrez/query.fcgi?cmd=Retrieve&db=PubMed&list_uids=12930561&dopt=Abstract)
[Abstract](http://www.biomedcentral.com/1471-2288/3/15/abstract)
Avialable in [html format](http://www.biomedcentral.com/1471-2288/3/15)
or [pdf format](http://www.biomedcentral.com/content/pdf/1471-2288-3-15.pdf)

+ The Ludbrook reference points out that laboratory
experiments often fail to account properly for early stopping
(interim analysis) and suggests an approach for handling this
properly.

You can find an [earlier version](http://www.pmean.com/04/EarlyStoppingAnimal.html) of this page on my [original website](http://www.pmean.com/original_site.html).