### 2021-10-23 text/21/zero-inflated-poisson.Rmd

---
title: "Zero-inflated models"
author: "Steve Simon"
source: "New"
date: "2021-09-20"
categories:
- Blog post
tags:
- Generalized linear model
output: html_document
---

In models of counts, zeros represent a special case. Often there is more than one reason why you might observe a zero count, and you might want to model this probability separately. One approach is known as a zero inflated Poisson model.

```{r setup}
suppressMessages(suppressWarnings(library(broom)))
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(pscl)))
```

### The zero inflated Poisson distribution

In any Poisson distribution, the probabilty of obtaining a zero is defined as

$P[Y=0]=\frac{\lambda^0e^{-\lambda}}{0!}=e^{-\lambda}$

Sometimes this probability is too low because there are multiple reasons why you might observe a zero. The zero-inflated Poisson model is a mixture of the Poisson random variable with a random variable that places all of its probability at zero.

$\pi \ 0 + (1-\pi)\frac{e^{-\lambda}\lambda^y}{y!}$

This changes the probability of zero to

$P[Y=0]=\pi+(1-\pi)e^{-\lambda}$

Since $\lambda$ has to be a positive value, the probability above is greater than the Poisson probability for any value of $\pi>0$.

The remaining probabilities are shrunk by a factor of $1-\pi$ to compensate for the increase in the probability at zero.

You can develop a separate regression model for $\pi$ the increase in the probability of zero in addition to the regression model for $\lambda$, the mean of the Poisson distribution.

The zero-inflated Poisson probability distribution is not part of the exponential family, so you need separate methods for estimation. You can find a function in the pscl library that will fit a zero-inflated Poisson model.

### Start with Poisson regression

It is easiest to understand the quasiPoisson model by comparing it to the Poisson model.

Francis Huang has a [nice web page][hua1] showing two datasets and how to analyze them using the Poisson regression model as well as several more sophisticated models. The first data set is available for anyone to download.

```{r}
fn <- "http://faculty.missouri.edu/huangf/data/poisson/articles.csv"
articles <- read.csv(fn)
head(articles)
```

The variables are
+ gender (fem),
+ marital status (mar),
+ number of children (kid5),
+ prestige of graduate program (phd), 
+ the number of articles published by the individual's mentor (ment)
+ the number of articles published by the scientist (art: the outcome).

To fit a Poisson regression model in R, use the glm function with the argument, family="poisson".

```{r}
pois1 <- glm(art ~ fem + ment + phd + mar + kid5, family=poisson, data = articles)
summary(pois1)
```

The interpretation of the coefficients should be done after exponentiating their values.

```{r}
pois1 %>%
  tidy %>%
  mutate(lower_limit=estimate-1.96*std.error) %>%
  mutate(upper_limit=estimate+1.96*std.error) %>%
  select(term, estimate, lower_limit, upper_limit) %>%
  mutate(estimate=exp(estimate)) %>%
  mutate(lower_limit=exp(lower_limit)) %>%
  mutate(upper_limit=exp(upper_limit)) -> pois1_coefficients
pois1_coefficients
```

### Fitting a zero-inflated-Poisson model

```{r}
zip1 <- zeroinfl(art ~ fem + ment + phd + mar + kid5 | 1, family=poisson, data = articles)
summary(zip1)
```

We cannot use the broom package to extract coefficients.

### References