### 2021-10-30 text/21/dark-side.Rmd
Started work on dark-side.Rmd

---
title: "The dark side of big data"
source: "New"
date: "2021-10-28"
categories:
- Blog post
tags:
- Big data
- Incomplete page
output: powerpoint_presentation
---

### Who am I?

Steve Simon

+ PhD Statistics, 1982, U Iowa
+ Teach in Biomedical and Health Informatics
  + Previous jobs at CMH, CDC
+ Part-time independent statistical consultant (P.Mean Consulting)
+ Married to a Pediatric Cardiologist (retired)
+ Run 5K and 4 mile races

<div class="notes">

Let me introduce myself. I am Steve Simon. I got a PhD in Statistics almost 40 years ago from the University of Iowa. The world has changed a lot since then, but I have tried to keep up. Today, if you want to sound trendy, you are a "data scientist". I teach in the Department of Biomedical and Health Informatics at UMKC. I have had previous jobs at Children's mercy Hospital, and the Centers for Disease Control and Prevention. I'm also a part-time statistical consultant. I have a sole proprietorship, P.Mean Consulting. That's short for Professor Mean. For people who don't get the joke, I point out that Professor Mean is not just your average Professor. Related to this talk, I should point out that I am obsessed with computers.

</div>

### Obsessed with computers since 1972

![Figure 1. Section on computer skills from my resume](http://www.pmean.com/new-images/21/dark-side-03.png)

### First poll question

![Figure 2. Quote from "Peggy Sue Got Married"](http://www.pmean.com/new-images/21/dark-side-01.png)

<div class="notes">

Think about your high school algebra class.

1. Do you remember any important formulas from that class?

2. Did you hate, hate, hate high school algebra?

3. Did you love high school algebra?

Big question: Will you use high school algebra in your future?

Source: https://www.moviequotes.com/s-movie/peggy-sue-got-married/

</div>

### Second poll question

![Figure 3. Images of various computers](http://www.pmean.com/new-images/21/dark-side-02.png)

<div class="notes">

How many of these computational devices have you used?

1. Laptop computer
2. Desktop computer
3. Tablet computer
4. Smart phone
5. Gaming console
6. Smart watch

Big question: What impact does the current generation's immersion of computing have on society?

Images found at

https://commons.wikimedia.org/wiki/File:Black_laptop_computer_open_frontal.svg
https://www.pcmag.com/picks/the-best-desktop-computers
https://www.cleverfiles.com/howto/what-is-tablet-computer.html
https://www.theverge.com/21420196/best-budget-smartphone-cheap
https://www.vulture.com/article/best-video-game-console-2020-ps5-xbox-series-x-nintendo-switch.html
https://www.homernews.com/marketplace/fitnus-smartwatch-review-legit-fitness-tracker-smart-watch/

</div>

### Are Statisticians Gods?

I'm helping someone who wants an alternative statistical analysis to the one used by the principal investigator. I'm happy to help and will offer advice about why my approach may be better, but I was warned that the PI considers the analysis chosen to be ordained by the **"Statistical Gods"** at her place of work.

<div class="notes">

True story. I was asked to review a report from a federal agency, but was warned that negative comments, even if accurate, might not be well received because the agency's work was ordained by their Statistic Gods. At first, I thought this was amusing. If I could get the title of "Statistical God", I could double my hourly consulting rate. But deep down this story really bothered me. It implies that Statistical skills are supernatural, and only available to a select few special people. I learned Statistics through hard work, and you can learn Statistics through hard work also. Some people will learn it faster than others because they have good aptitudes in mathematics and programming. But there is nothing other than time to stop you.

</div>

### Statistics are a social construct

![Figure 4. Cover from Joel Best's book](http://www.pmean.com/new-images/21/dark-side-04.png)

<div class="notes">

I'm going to share a message that is controversial, but which I believe deeply in. It's a message from a book by Joel Best: Lies, Damned Lies, and Statistics. Untangling Numbers from the Media, Politicians, and Activists. He argues that Statistics are a social construct. I need to define social construct here. A social construct is a method by which people create their perception of reality.

That's certainly true of Statistics. Statistcs are "selected, shaped, and presented by human beings"

Source of quote: Milo Shield, Teaching the Social Construction of Statistics.  Available at https://www.researchgate.net/publication/277299441_Teaching_the_Social_Construction_of_Statistics

Image taken from https://www.ucpress.edu/book/9780520274709/damned-lies-and-statistics.

</div>

### The arrogance of empiricism (1/2)

"I often say that when you can measure what you are speaking about, and express it in numbers, you know something about it; but when you cannot measure it, when you cannot express it in numbers, your knowledge is of a meagre and unsatisfactory kind; it may be the beginning of knowledge, but you have scarcely in your thoughts advanced to the state of Science, whatever the matter may be." Lord Kelvin.

<div class="notes">

The idea that Statistics are a social construction flies in the face of empiricism, the philosophy that experiments can reveal the realities of the world.

Here is a quote in support of empiricism. I won't read all of the first quote, but notice the sneering phrase that without numbers, "your knowledge is of a meagre and unsatisfactory and unsatisfactory kind."

</div>

### The arrogance of empiricism (2/2)

"No human investigation can be called real science if it cannot be demonstrated mathematically." Leonardo da Vinci

<div class="notes">

Here is a second quote that is just as bad. Science without numbers (math) is not real.

</div>

### Why empiricism fails (1/3)

"The government is very keen on amassing statistics. They collect them, add them, raise them to the nth power, take the cube root and prepare wonderful diagrams. But you must never forget that every one of these figures comes in the first instance from the village watchman, who just puts down what he damn well pleases." Sir Josiah Stamp

<div class="notes">

This quote from Josiah Stamp illustrates why Statistics are a social construct. It doesn't matter how fancy we dress up the numbers, nth powers, cube roots. It all comes down to the village watchman who does what he damn well pleases.

The village watchman is part of the social constuction of Statistics. So are the government officials who do all sorts of mathematical manpulations which provide an impression of objectivity.

</div>

### Why empricism fails (2/3)

![Figure 5. Cover of Stephen Jay Gould's book](http://www.pmean.com/new-images/21/dark-side-05.png)

<div class="notes">

There's a great book by Stephen Jay Gould that really makes you think hard about numbers. The book, The Mismeasure of Man, talks about IQ testing from a historical perspective and totally destroys the idea that intelligence is a trair that can be easily quantified. He starts with efforts in the 1800s to use the size of a person's skull as a measure of their intelligence. A big skull means a big brain that lives inside there and bigger brains are smarter brains.

</div>

### Why empricism fails (3/3)

![Figure 6. Frame from Yes, Prime Minister video clip](http://www.pmean.com/new-images/19/leading-questions01.png)

<div class="notes">

This short video clip is an excellent illustration of how the questions leading up to a particular question on a survey can bias the response to that survey. It comes from a British comedy, Yes Prime Minister, that ran in the 1980s.

YouTube. Leading Questions- Yes Prime Minister. Taken from the 1st Season of Yes Prime Minister – Episode 2, The Ministerial Broadcast. 2:16 running time. Published on Jan 15, 2012.

https://www.youtube.com/watch?v=G0ZZJXw4MTA

</div>

### But numbers still have value

Recognizing that Statistics are a social construction does not mean that Statistics are worthless and that experiments are a waste of time.

### Example: Quality of Life

![Figure 7. First few questions from SF-36 form](http://www.pmean.com/new-images/21/dark-side-08.png)

<div class="notes">

Quality of Life is a measure that is reported in many research studies. There is no question that quality of life is a social construction. It is measured by forms like the SF-36 shown here. But it is not an objective reality but something that was constructed by humans.

Now I mentioned earlier that I like to run. It is a pretty strenuous activity, but one that I enjoy. A friend of mine said, though, that she couldn't understand why you would want to run, unless you were being chased. One the other hand, I would not miss a thing if I did not push a vacuum, bowl, or play golf.

There is no objective measure of quality of life. It depends on the values of the people who are developing forms like the SF-36.

But we can still use social constructs like quality of life, as long as we do it carefully.

Image obtained from

https://clinmedjournals.org/articles/jmdt/jmdt-2-023-figure-1.pdf

</div>

### Rules for using Statistics as social constructs

1. Understand the context in which the Statistic was generated.

2. Identify possible biases

3. Recognize limitations

4. Beware of confirmation bias

5. Avoid nihilistic thinking

<div class="notes">

</div>

### Things have gotten worse, thanks to big data

1. Large amounts of data of uneven quality

2. Too much trust of objectivity of numbers

2. Black box models

3. Lack of accountability

4. Scaling problems

5. Loss of privacy

<div class="notes">

</div>

### Weapons of Math Destruction

![Figure 6. Cover of Cathy O'Neill's book](http://www.pmean.com/new-images/21/dark-side-06.png)

<div class="notes">

</div>

### Myths about big data

1. Algorithms are objective

2. If you have enough data, quality is no longer an issue

3. We are getting better at this

### No single cause of these problems

1. wrong data

2. Wrong objective

3. Wrong deployment

4. Wrong team

### What they say data science is

```{r setup, echo=FALSE}
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

```{r, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.6
v <- 0.85
draw_circles <- function() {
  dx <- 1
  dy <- 0.866
  i <- seq(0, 2*pi, length=100)
  circle <- data.frame(x=sin(i), y=cos(i))
  circle %>%
    mutate(g=1) %>%
    mutate(x=x-o*dx) %>%
    mutate(y=y-o*dy) -> circle1
  circle %>%
    mutate(g=2) %>%
    mutate(x=x+o*dx) %>%
    mutate(y=y-o*dy) -> circle2
  circle %>%
    mutate(g=3) %>%
    mutate(y=y+o*dy) -> circle3

  circle1 %>%
    bind_rows(circle2) %>%
    bind_rows(circle3) %>%
    ggplot(aes(x, y, group=g)) + 
      geom_path(color=rep(c("red", "darkgreen", "blue"), each=100), size=3) +
      geom_text(x=v*dx, y=-v*dy, size=5, label="Math,\nStat", color="darkgreen") +
      geom_text(x=-v*dx, y=-v*dy, size=5, label="Coding,\nComputers", color="red") +
      geom_text(x=0, y=v*dy, size=5, label="Science,\nMedicine", color="blue") +
      expand_limits(x=c(-(dx+1),(dx+1))) +
      expand_limits(y=c(-(dx+1),(dx+1))) +
      theme_void()
}
o <- 0.6
v <- 0.85
draw_circles() + geom_text(x=0, y=-0.1, label="Data\nScience", size=5)
```

<div class="notes">

</div>

### What data science really is

```{r, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.88
draw_circles() + geom_text(x=0, y=-0.1, label="Data\nScience", size=10)
```

<div class="notes">

</div>

### Why you are needed

1. Too many geeks, not enough scientists
2. More racial, gender diversity is needed

If you are interested, email me: simons@umkc.edu

<div class="notes">

</div>

### Readings

[sim1]: http://new.pmean.com/amazon-bias/
[sim2]: http://new.pmean.com/big-data-bs/
[sim3]: http://new.pmean.com/big-data-goes-to-school/
[sim4]: http://new.pmean.com/lending-false-credibility/
[sim5]: http://new.pmean.com/ai-story/