### 2021-11-12 text/15/missing-values-in-r.Rmd

---
title: "PMean: Missing values in R talk"
author: "Steve Simon"
source: "http://blog.pmean.com/missing-values-in-r/"
date: "2015-02-07"
category: Blog post
tags: R software
output: html_document
---

I'm talking a bit about missing values in R this afternoon for the
Kansas City R Users Group. Here is what I'll be talking
about.

<!---More--->

When you import data into R, the first thing you should do is

KNOW YOUR COUNT

This means that you need to know how many rows are in your data, and for
all of the key variables, how many missing values you have. In a nested
or hierarchical model, you need to also know the counts at each level of
nesting or each level of the hierarchy.

Here is a data set, freely available on the web, that shows sleep
patterns and other statistics on various mammals. It has a handful of
missing values and can help illustrate how you might tackle a new
dataset. You can find a description of this data set in [html format][sta1].

```{r}
sl <- read.table(file="http://www.statsci.org/data/general/sleep.txt",header=TRUE)
dim(sl)
```

```{r}
names(sl)
```

This data set has 62 rows and 11 variables. Since the data set doesn't
have too many variables, let's print out the first few and last few
rows.

```{r}
head(sl)
```

```{r}
tail(sl)
```

The summary function is useful, as it will display percentiles and the
mean for numeric variables, but also (very importantly) display the
number of missing values.

For factors, it will display a table of the first seven most frequently
occuring values

```{r}
summary(sl)
```

The first variable, Species, has 62 unique values. You can tell this by
the fact that the first seven most frequently occurring values all had
frequency 1. There are several other ways to check for uniqueness.

```{r}
table(table(sl$Species))
```

```{r}
length(unique(sl$Species))
```

    ## [1] 62

The table function does not show NAs, at least with the default option

```{r}
table(sl$Gestation)
```

You would notice this if you added a total to the end of the table.

```{r}
addmargins(table(sl$Gestation))
```

The total is only 58, which is less than the magic number of 62. This
means that there are 4 missing values.

You can also force the table to show counts for missing values.

```{r}
table(sl$Gestation,useNA="always")
```

The default option for useNA is "never" and a third option is "ifany".
Notice how "ifany" differs from "always"

```{r}
table(sl$Gestation,useNA="ifany")
```

```{r}
table(sl$Danger,useNA="ifany")
```

```{r}
table(sl$Danger,useNA="always")
```

Many other functions in R offer multiple ways to handle NAs.

```{r}
mean(sl$Gestation,na.rm=FALSE)
```

```{r}
mean(sl$Gestation,na.rm=TRUE)
```

```{r}
mod1 <- lm(TotalSleep~Gestation,data=sl,na.action=na.omit)
summary(mod1)
```

The alternate option for na.action in the lm function is na.fail, but I
cannot run this directly in knitr, because the resulting error message
causes knitr to stop.

Logic values involving NAs are a bit tricky. You cannot get at NAs
directly using the == comparison operator.

```{r}
which(sl$Gestation==NA)
```

    ## integer(0)

What's happening here is that most logic comparisons invloving NA will
evaluate not to TRUE or FALSE, but rather to NA.

If you want to select missing values, you must use the is.na function.

```{r}
which(is.na(sl$Gestation))
```

    ## [1] 13 19 20 56

The is.finite function represents "opposite" function, though there are
a few subtle differences.

```{r}
which(is.finite(sl$Gestation))
```

    ##  [1]  1  2  3  4  5  6  7  8  9 10 11 12 14 15 16 17 18 21 22 23 24 25 26
    ## [24] 27 28 29 30 31 32 33 34 35 36 37 38 39 40 41 42 43 44 45 46 47 48 49
    ## [47] 50 51 52 53 54 55 57 58 59 60 61 62

Logic involving missing values can seem at times a bit illogical.

```{r}
sl$Gestation==NA
```

    ##  [1] NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA
    ## [24] NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA
    ## [47] NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA NA

```{r}
sl$Gestation==645
```

    ##  [1]  TRUE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE
    ## [12] FALSE    NA FALSE FALSE FALSE FALSE FALSE    NA    NA FALSE FALSE
    ## [23] FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE
    ## [34] FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE
    ## [45] FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE FALSE
    ## [56]    NA FALSE FALSE FALSE FALSE FALSE FALSE

```{r}
tst.na <- NA
tst.na==tst.na
```

    ## [1] NA

Why would NA==NA not evaluate to TRUE? You could argue that a missing
value might be 3 or 7 or 923.427, so you would never know if it equaled
something else, even another missing value.

This produces a three valued logic system.

```{r}
tst.tvl <- c(TRUE,FALSE,NA)
not.table <- !tst.tvl
names(not.table) <- tst.tvl
not.table
```

    ##  TRUE FALSE  <NA> 
    ## FALSE  TRUE    NA

```{r}
and.table <- outer(tst.tvl,tst.tvl,"&")
dimnames(and.table) <- list(tst.tvl,tst.tvl)
and.table
```

    ##        TRUE FALSE  <NA>
    ## TRUE   TRUE FALSE    NA
    ## FALSE FALSE FALSE FALSE
    ## <NA>     NA FALSE    NA

```{r}
or.table <- outer(tst.tvl,tst.tvl,"|")
dimnames(or.table) <- list(tst.tvl,tst.tvl)
or.table
```

Other programs, such as SAS, use a two valued logic system. What SAS
does is to set missing values equal to negative infinity. This seems
simpler because it allows you to keep a two-valued logic system. But it
has its own problems.

Note how R handles the less than comparison operator with missing
values.

```{r}
sl$Gestation<50
```

In SAS, this comparison would evaluate to TRUE for the four missing
values.

The only way to select missing values is with the is.na function.

There are three more values in R that are closely related to missing.

```{r}
1/0
```

```{r}
-1/0
```

```{r}
sqrt(-1)
```

The difference between Inf and NaN is in how comparisons are made.

```{r}
Inf>3
```

```{r}
NaN>3
```

```{r}
Inf==Inf
```

```{r}
NaN==NaN
```

R allows you to make a few computations using Inf, but others are not
allowed.

```{r}
Inf+1
```

```{r}
1/Inf
```

```{r}
0*Inf
```

```{r}
Inf-Inf
```

It's also worth noting that you can force R to use complex numbers for
the square root of -1, although this is not the default.

```{r}
sqrt(as.complex(-1))
```

The is.na and is.finite functions are not perfect opposites.

```{r}
tst.cases <- c(3,NA,NaN,Inf,-Inf)
is.na.table <- is.na(tst.cases)
names(is.na.table) <- tst.cases
is.na.table
```

```{r}
is.finite.table <- is.finite(tst.cases)
names(is.finite.table) <- tst.cases
is.finite.table
```

You can find an [earlier version][sim1] of this page on my [blog][sim2].

[sim1]: http://blog.pmean.com/missing-values-in-r/
[sim2]: http://blog.pmean.com

[sta1]: http://www.statsci.org/data/general/sleep.html