### 2021-12-12 xxtemp.Rmd

---
title: "Two page example for Powerpoint"
author: "Steve Simon"
date: "12/11/2021"
output: powerpoint_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

### R Markdown

This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with Bullets

- Bullet 1
- Bullet 2
- Bullet 3

## Slide with R Output

```{r cars, echo = TRUE}
summary(cars)
```

## Slide with Plot

```{r pressure, echo=FALSE}
plot(pressure)
```

## test

```{r}
plot(1:4, 2:5)
```

## Column format

:::::: {.columns}
::: {.column}
Content of the left column.

```{r}
rnorm(10)
```
:::

::: {.column}
Content of the right column.

```{r, echo=FALSE}
plot(1:4, 2:5)
```

:::
::::::