### 2022-01-07 text/22/mapping-longitude-latitude.Rmd
Started work on mapping-longitude-latitude.Rmd

---
title: "Mapping latitude and longitude"
author: "Steve Simon"
source: new
date: "2022-01-06"
category: 
- Blog post
tags:
- Geocoding
output: html_document
---

I'm trying to understand some of the basics of mapping and found an interesting example on [this website]

```{r usa-map}
library(maps); library(ggplot2); library(mapproj)
states <- map_data("state")
usamap <- ggplot(states, aes(x=long, y=lat, group=group)) +
  geom_polygon(fill="white", colour="black")
usamap + coord_map("mercator")
usamap + coord_map("azequalarea") 
```

It helps to peek at the first few rows of data.

```{r peek}
head(states)
```

The plot is based on longitude/latitude pairs. The first thing to realize is that a direct plot without the coor_map function fails spectacularly.

```{r epic-fail}
usamap <- ggplot(states, aes(x=long, y=lat, group=group)) +
  geom_polygon(fill="white", colour="black")
```

Additional references are [here][ucs1], [here][ear1], [here][mgi1], and [here][tid1].

[ucs1]: https://www.nceas.ucsb.edu/sites/default/files/2020-04/OverviewCoordinateReferenceSystems.pdf
[ear1]: https://www.earthdatascience.org/courses/earth-analytics/spatial-data-r/understand-epsg-wkt-and-other-crs-definition-file-types/
[mgi1]: https://mgimond.github.io/Spatial/coordinate-systems-in-r.html
[tid1]: https://ggplot2.tidyverse.org/reference/coord_map.html
[rpu1]: https://rpubs.com/Augusto86/697809