### 2022-07-06 text/22/different-parameterizations.Rmd
Started work on different-parameterizations.md

---
title: "Different parameterizations for one factor ANOVA"
author: "Steve Simon"
source: new
date: "2022-07-06"
category: 
- Blog post
tags:
- Analysis of variance
output: html_document
---

I'm trying to understand some of the basics of contrasts in analysis of variance. In particular, I want to figure out some alternative parameterizations of the coefficients that are estimated as part of the ANOVA model.

```{r setup}
library(MASS)
library(matlib)
library(tidyverse)
knitr::opts_chunk$set(echo = TRUE)
```

First, let's set up an artificial dataset that is nice and easy to work with.

```{r}
means <- c(50, 52, 60, 70)
mean(means)
y <- 
  rep(means, each=3) +
  rep(c(0, 1, -1), 4)
g <- rep(1:4, each=3)
df <- cbind(g, y)
```

Notice the pattern in the group means.

```{r}
data.frame(g, y) %>%
  group_by(g) %>%
  summarize(mean_y=mean(y))
```

I want to use the lm function to fit the data, because it is easier to peek at the underlying coefficients.

### The simplest parameterization

The simplest parameterization is to let each beta coefficient represent a different mean. We have to drop the intercept term for this model to work properly.

```{r}
x1 <- as.numeric(g==1)
x2 <- as.numeric(g==2)
x3 <- as.numeric(g==3)
x4 <- as.numeric(g==4)
```

This is what the X matrix looks like.

```{r}
x <- cbind(x1, x2, x3, x4) 
x
```

This is what the regression results look like.

```{r}
lm(y~x1+x2+x3+x4-1)
```

It may help to view the underlying matrix calculations. Recall that

$\hat{\beta}=(X'X)^{-1}X'Y$

So the matrix

$\hat{\beta}=(X'X)^{-1}X'$

will tell you how each parameter in the regression model is a linear combination of the Y's.

```{r}
inv(t(x)%*%x)%*%t(x) %>% fractions
```

So the first coefficient is an average of the first three y values, the second is an average of the next three y values, etc.

### Indicator variables

A common way to parameterize the ANOVA model is to use an intercept term and indicator variables for all but one of the groups.

```{r}
x0 <- rep(1, 12)
x1 <- as.numeric(g==2)
x2 <- as.numeric(g==3)
x3 <- as.numeric(g==4)

```

This is what the X matrix looks like.

```{r}
x <- cbind(x0, x1, x2, x3) 
x
```

This is what the regression results look like.

```{r}
lm(y~x0+x1+x2+x3-1)
```

```{r}
inv(t(x)%*%x)%*%t(x) %>% fractions
```

So the first coefficient is an average of the first three y values (Y1, Y2, Y3), the second is the difference between the average of the second three y values (Y4, Y5, Y6) and the average of the first three y values (Y1, Y2, Y3), etc.

This seems counter intuitive. The slope coefficient for $\beta_2$ involves a column with zeros everywhere except for some ones for Y4, Y5, and Y6. How could that coefficient end up involving Y1, Y2, and Y3?

The point to remember is that the slope coefficient in a linear regression with multiple independent variables has a slightly more complex interpretation than the slope coefficient in a linear regression with a single independent variable. In the later case, the interpretation would be

+ the estimated average change in $Y$ when $X$ increases by one unit.

In the former case, the interpretation would be

+ the estimated average change in $Y$ when $X_i$ increases by one unit **while holding all the other** $X$**'s constant**.

So the slope associated with X1 requires you to hold X2 and X3 constant. You can hold them constant at 0 or hold them constant at 1, but the former is simpler to understand. THink of it as moving from one of the first three rows of your X matrix.

```{r}
x[1:3, ]
```

to one of the next three rows of your X matrix

```{r}
x[4:6, ]
```

This is effectively a change from the mean of the first three observations (Y1, Y2, Y3) subtracted from the mean of the next three observations (Y4, Y5, Y6).

In writing this webpage, I relied on a very helpful tutorial paper:

Daniel J. Schad, Shravan Vasishth, Sven Hohenstein, Reinhold Kliegl. How to capitalize on a priori contrasts in linear (mixed) models: A tutorial. Journal of Memory and Language, 2020, 110. Available in [html format][sch1] or [pdf format][sch2].

as well as a page on the UCLA Statistics site

R Library Contrast Coding Systems for Categorical Variables. Available in [html format][ucla1]

[ucla1]: https://stats.oarc.ucla.edu/r/library/r-library-contrast-coding-systems-for-categorical-variables/

[sch1]: https://doi.org/10.1016/j.jml.2019.104038
[sch2]: https://www.sciencedirect.com/science/article/pii/S0749596X19300695/pdfft?md5=05f53f6c3386f0e0c6ae203178b33e33&pid=1-s2.0-S0749596X19300695-main.pdf