### 2022-12-06 text/22/definition-schoenfeld-residuals.Rmd

---
title: "Schoenfeld residuals"
author: "Steve Simon"
date: '2022-11-30'
output:
  html_document: default
category: Blog post
tags: Survival analysis
source: new
---

There are several different types of residuals in a Cox proportional hazards model. One of them is the Schoenfeld residuals. They are used to test the proportional hazards assumption.

<!---more--->

To understand the Schoenfeld residuals, you need to first recall the partial likelihood that is used to estimate hazard ratios in Cox regression model.

$L_i(\beta)  
 =\frac{\lambda(Y_i\mid X_i)}{\sum_{j:Y_j\ge Y_i}\lambda(Y_i\mid X_j)} 
 =\frac{\lambda_0(Y_i)\theta_i}{\sum_{j:Y_j\ge Y_i}\lambda_0(Y_i)\theta_j}
 =\frac{\theta_i}{\sum_{j:Y_j\ge Y_i}\theta_j},$

where

$\theta_j = exp(X_j \beta)$