### 2022-12-06 text/22/definition-schoenfeld-residuals.Rmd
Added comments about likelihood in definition-schoenfeld-residuals.

---
title: "Schoenfeld residuals"
author: "Steve Simon"
date: '2022-11-30'
output:
  html_document: default
category: Blog post
tags: Survival analysis
source: new
---

There are several different types of residuals in a Cox proportional hazards model. One of them is the Schoenfeld residuals. They are used to test the proportional hazards assumption.

<!---more--->

To understand the Schoenfeld residuals, you need to first recall the partial likelihood that is used to estimate hazard ratios in Cox regression model.

The likelihood function is the foundation for most statistical methods. You are trying to estimate some unknown parameters $\theta$ and each subject has a density function that is related to data measured on that subject and the unknown parameters.

Here's a simple example. Assume that there is no censoring and that the event time is exponential with a scale parameter $\lambda_0$ for the control group and a mean parameter that is $\lambda_0+\lambda_1$ for the exposed or treatment group. The data in this case is the actual event time Y_i and an indicator variable X_i that is equal to zero for the control group and 1 for the exposed group.

To simplify things, I will let

$\lambda=(\lambda_0, \lambda_1)$

$Y=(Y_1, Y_2, ..., Y_n)$

and define X similarly.

The density function,evaluated at the observed survival time is

$f_i(Y_i, X_i, \lambda)=(\lambda_0+\lambda_1X_i)exp(-Y_i(\lambda_0+\lambda_1X_i))$

We want to find "good values" of $\lambda$, ones that are associated with large (and therefore more likely) values of the density. The maximum likelihood estimate is the value of $lambda$ that maximizes the products of all these densities. For the exponential case, this would be

$L(\lambda, Y, X)=\prod_i (\lambda_0+\lambda_1X_i) exp(-Y_i/(\lambda_0+\lambda_1X_i))$

Let's make up some simple data for this example. Let Y=(1, 5, 2) and X=(0, 0, 1). The likelihood function in this simple case is

$L(\lambda, Y, X)=\lambda_0 exp(-1 \lambda_0) \lambda_0 exp(-5 \lambda_0) (\lambda_0+\lambda_1) exp(-5(\lambda_0+\lambda_1))$

It is often simpler to maximize the log of the likelihood. There are several reasons for this. First, a product of densities turns into a sum of log densities and sums are easier to work with than products. Second, many densities have an exponential function built into them and that function "disappears" after taking logs. I will use lower case for the log likelihood.

$l(\lambda, Y, X)=log\lambda_0 - \lambda_0 + log\lambda_0 - 5 \lambda_0 + log(\lambda_0 + \lambda_1) -2 (\lambda_0+\lambda_1))$

or

$l(\lambda, Y, X) = 2 log\lambda_0 -9 \lambda_0 + log(\lambda_0+\lambda_1) -2 \lambda_1$

Let's graph this function.

```{r}
n0 <- 20
n1 <- 21

u0 <- seq(1/n0, 0.7, length=n0)
u1 <- seq(0, 1.4, length=n1)

l <- function(u0, u1) {
  2*log(u0)-9*u0+log(u0+u1)-2*u1
}

l_matrix <- matrix(nrow=n0, ncol=n1)

for (i0 in 1:n0) {
  for (i1 in 1:n1) {
    l_matrix[i0, i1] <- l(u0[i0], u1[i1])
  }
}

plot(u0, l_matrix[1:n0, 1])
plot(u1, l_matrix[1, 1:n1])
persp(u0, u1, l_matrix, theta=15, phi=30)
```

$L_i(\beta)  
 =\frac{\lambda(Y_i\mid X_i)}{\sum_{j:Y_j\ge Y_i}\lambda(Y_i\mid X_j)} 
 =\frac{\lambda_0(Y_i)\theta_i}{\sum_{j:Y_j\ge Y_i}\lambda_0(Y_i)\theta_j}
 =\frac{\theta_i}{\sum_{j:Y_j\ge Y_i}\theta_j},$

where

$\theta_j = exp(X_j \beta)$