### 2023-01-06 text/22/2022.Rmd
Added 2023.md

---
title: "Pages written in 2022"
source: "New"
date: "2022-12-31"
categories:
- Blog post
tags:
- Website details
output: html_document
---

```{r setup, echo=FALSE}
library(glue)
library(magrittr)
library(stringr)
yr <- 2023:1999
glue("[{yr}](http://new.pmean.com/{yr}/)") %>%
  paste(collapse=" | ") -> yr_list
list.files() %>%
  str_subset("md$") %>%
  str_replace(fixed(".Rmd"), "") %>%
  str_replace(fixed(".md"), "") -> f
n <- length(f):1
header <- glue("New website ({n[1]} files)")
glue("+ [{f}](http://new.pmean.com/{f}/)  ") %>%
  paste(collapse="  \n") -> file_list
```

`r yr_list`

`r header`

`r file_list`