### 2023-02-21 text/22/box-with-no-whiskers.Rmd

---
title: "My boxplot shows a box but no whiskers"
author: "Steve Simon"
date: '2022-12-15'
output:
  html_document: default
category: Blog post
tags:
- Graphical display
source: new
---

One of my favorite graphs is the box plot. It is a visualization of a five number summary. Those five numbers are minimum, 25th percentile, median (aka the 50th percentile), 75th percentile, and maximum. The box in a box plot extends from the 25th to the 75th percentile and a line inside the box represents the median.

It's great for comparing two groups. Here's a boxplot showing two groups. Notice that the first group tends to have larger values.

```{r, fig.height=1.5, echo=FALSE}
suppressMessages(suppressWarnings(library(tidyverse)))
n1 <- 10
n2 <- 10
pseudo_boxplot <- function(n1, n2, a1, b1, a2, b2) {
  x1 <- qbeta((1:n1)/(n1+1), 6, 2)
  x2 <- qbeta((1:n2)/(n2+1), 5, 3)
  data.frame(x=c(x1, x2), g=paste0("g", rep(1:2, c(n1, n2)))) -> df
  print(df)
  df %>%
    ggplot(aes(g, x)) +
      geom_boxplot() +
      coord_flip()
}

pseudo_boxplot(10, 10, 6, 2, 5, 3)
```

What happens if you try to compute the five number summary with a dataset that only has only two values?

```{r, fig.height=1.5, echo=FALSE}
pseudo_boxplot(10, 2, 6, 2, 5, 3)
```