### 2023-02-28 text/23/gamma-hedging-hyperprior.Rmd

---
title: "The hedging hyperprior for the exponential distribution"
source: "New"
date: "2023-02-28"
categories:
- Blog post
tags:
- Bayesian statistics
output: html_document
---

I want to write a paper about the hedging hyperprior. I need to work out some simple examples first. Here is an example involving the gamma distribution.

<!---more--->

Consider a Bayesian analysis of an outcome variable that has an exponential distribution. The density of the gamma distribution is

$f(x) = \theta^{-1} e^{-x / \theta}$

You have an informative prior on $\theta$ that has an inverse gamma distribution

$g(\theta; \alpha, \beta)=\frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{-\alpha-1}e^{-\beta / \theta}$

and you collect a sample of n data points, $\mathbf{X} = X_1, X_2, ..., X_n$. This gives you a likelihood of

$\prod_{i=1}^n f(X_i)=\theta^{-n} e^{-n \bar X / \theta}$

and a posterior distribution of

$h(\theta | \mathbf{X}) \propto \frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{-\alpha-1}e^{-\beta / \theta}\theta^{-n} e^{-n \bar X / \theta}$

which simplifies to

$h(\theta | \mathbf{X}) \propto \theta^{-\alpha-1-n}e^{-(\beta + n\bar X) / \theta}$

which, thanks to the miracle of conjugancy, is an inverse gamma distribution. The posterior mean

$\frac{\beta + n\bar X}{\alpha-1+n}$

can be written as a weighted average of the prior mean and the mean of the observed data.

$\Big(\frac{\alpha-1}{\alpha-1+n}\Big)\frac{\beta}{\alpha-1}+\Big(\frac{n}{\alpha-1+n}\Big)\bar X$

Let's illustrate this with an informative prior with $\alpha$ = 10 and $\beta$ = 27.

This corresponds to a fairly informative prior distribution. This distribution places $\theta$ within `r sprintf("%.2f", 1/qgamma(0.95, 10, 27))` and `r sprintf("%.2f", 1/qgamma(0.05, 10, 27))` with 95% probability. If you observe 30 values of x with a mean of 7.7.