### 2023-03-06 text/23/gamma-hedging-hyperprior.Rmd
Added contour plot to gamma-hedging-hyperprior

---
title: "The hedging hyperprior for the exponential distribution"
source: "New"
date: "2023-02-28"
categories:
- Blog post
tags:
- Bayesian statistics
output: html_document
---

```{r setup}
suppressMessages(suppressWarnings(library(MASS)))
suppressMessages(suppressWarnings(library(tidyverse)))
knitr::opts_chunk$set(echo=TRUE, fig.width=8, fig.height=2)
```

I want to write a paper about the hedging hyperprior. I need to work out some simple examples first. Here is an example involving the gamma distribution.

<!---more--->

### A simple Baysian analysis with an informative prior.

Consider a Bayesian analysis of an outcome variable that has an exponential distribution. The density of the gamma distribution is

$f(x) = \theta^{-1} e^{-x / \theta}$

You have an informative prior on $\theta$ that has an inverse gamma distribution

$g(\theta; \alpha, \beta)=\frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{-\alpha-1}e^{-\beta / \theta}$

and you collect a sample of n data points, $\mathbf{X} = X_1, X_2, ..., X_n$. This gives you a likelihood of

$L(\mathbf{X})=\prod_{i=1}^n f(X_i)=\theta^{-n} e^{-n \bar X / \theta}$

and a posterior distribution of

$h(\theta | \mathbf{X}) \propto \frac{\beta^\alpha}{\Gamma(\alpha)}\theta^{-\alpha-1}e^{-\beta / \theta}\theta^{-n} e^{-n \bar X / \theta}$

which simplifies to

$h(\theta | \mathbf{X}) \propto \theta^{-\alpha-1-n}e^{-(\beta + n\bar X) / \theta}$

which, thanks to the miracle of conjugancy, is an inverse gamma distribution. The posterior mean

$\frac{\beta + n\bar X}{\alpha-1+n}$

can be written as a weighted average of the prior mean and the mean of the observed data.

$\Big(\frac{\alpha-1}{\alpha-1+n}\Big)\frac{\beta}{\alpha-1}+\Big(\frac{n}{\alpha-1+n}\Big)\bar X$

```{r}
x_max <- 3

a0 <- 41
b0 <- 44

n <- 60
xbar <- 2.1

a1 <- a0 + n
b1 <- b0 + n*xbar

posterior_alpha <- a0+n
posterior_beta <- b0+n*xbar

x_grid <- seq(0.01, x_max, length=1000)

f <- function(x, theta) {
  theta^(-1) * 
    exp(-x/theta)
}

g <- function(theta, alpha, beta) {
  beta^alpha/gamma(alpha) *
    theta^(-alpha-1) * 
    exp(-beta/theta)
}

prior05 <- 1/qgamma(c(0.975), a0, b0) 
prior95 <- 1/qgamma(c(0.025), a0, b0)

post05 <- 1/qgamma(c(0.975), a1, b1) 
post95 <- 1/qgamma(c(0.025), a1, b1)

```

Let's illustrate this with an informative prior with $\alpha$ = `r a0` and $\beta$ = `r b0`.

```{r}
g_grid <- g(x_grid, a0, b0)
sum(g_grid)*0.01 # Check the g is a density.
outside <- x_grid < prior05 | x_grid > prior95
x_color <- 
  ifelse(outside, "white", "gray")
data.frame(x=x_grid, y=g_grid) %>%
  ggplot(aes(x, y)) +
    geom_segment(aes(xend=x, yend=0), color=x_color) +
    geom_line() + xlab(" ") + ylab(" ")
```

This corresponds to a fairly informative prior distribution. This distribution places $\theta$ within `r sprintf("%.2f", prior05)` and `r sprintf("%.2f", prior95)` with 95% probability.

Let's see what happens with this informative prior if you observe `r n` values of x with a mean of `r xbar`. The likelihood is

```{r}
f_grid <- f(xbar, x_grid)^n
x_color <- "white"
data.frame(x=x_grid, y=f_grid) %>%
  ggplot(aes(x, y)) +
    geom_line() + xlab(" ") + ylab(" ")
```

Multiply these two together to get the posterior distribution.

```{r}
h_grid <- g(x_grid, a1, b1)
outside <- x_grid < post05 | x_grid > post95
x_color <- 
  ifelse(outside, "white", "gray")
data.frame(x=x_grid, y=h_grid) %>%
  ggplot(aes(x, y)) +
    geom_segment(aes(xend=x_grid, yend=0), color=x_color) +
    geom_line() + xlab(" ") + ylab(" ")
```

The posterior distribution is not close to either the prior distribution or the likelihood. It places $\theta$ within `r sprintf("%.2f", post05)` and `r sprintf("%.2f", post95)` with 95% probability. It excludes the prior mean of `r sprintf("%.2f", b0/(a0-1))` and it excludes `r xbar`, the mean of the data.

### The hedging hyperprior

When the likelihood of your data and your informative prior are in conflict, you may wish to downweight the informative prior. You can do this by including a hedging hyperprior.

Define your prior distribution as an inverse gamma distributions with parameters $1+\lambda (\alpha-1)$ and $\lambda \beta$ where $\lambda$ ranges between 0 and 1. Then place a hyperprior on the parameter $\lambda$. The uniform(0,1) distribution is a reasonable choice for the hyperprior, but others are possible.

```{r}
knitr::opts_chunk$set(echo=TRUE, fig.width=8, fig.height=8)
```

```{r}
tau_grid <- seq(0, 1, length=60)
theta_grid <- seq(0, x_max, length=60)

gstar <- function(theta, tau) {g(theta, tau*(a0-1)+1, tau*(b0-1)+1)}
density <- outer(theta_grid, tau_grid, gstar)
op <- par(bg = "white")
z_max <- max(density, na.rm=TRUE)
persp(
  theta_grid, 
  tau_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="theta",
  ylab="tau",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

It is hard to see at this scale, but the prior distribution is higher for small values of $\tau$ when $\theta4 is very large or very small. This becomes more obvious if you remove the middle of the surface and rescale the Z-axis.

```{r}
density <- outer(theta_grid, tau_grid, gstar)
density[11:40, ] <- NA
op <- par(bg = "white")
z_max <- max(density, na.rm=TRUE)
persp(
  theta_grid, 
  tau_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="theta",
  ylab="tau",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

The likelihood function is constant across all values of $\tau$, so its surface looks like the following.

```{r}
tau_grid <- seq(0, 1, length=60)
theta_grid <- seq(0, x_max, length=60)

fstar <- function(theta, tau) {f(xbar, theta)^n}
likelihood <- outer(theta_grid, tau_grid, fstar)
z_max <- max(likelihood, na.rm=TRUE)
persp(
  theta_grid, 
  tau_grid, 
  likelihood, 
  zlim=c(0, 1.1*z_max),
  xlab="theta",
  ylab="tau",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

The product of the prior and the likelihood is proportional to the posterior distribution.

```{r}
tau_grid <- seq(0, 1, length=60)
theta_grid <- seq(0, x_max, length=60)

hstar <- function(theta, tau) {g(theta, tau*(a0-1)+1, tau*(b0-1)+1)*f(xbar, theta)^n}
density <- outer(theta_grid, tau_grid, hstar)
z_max <- max(density, na.rm=TRUE)
persp(
  theta_grid, 
  tau_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="theta",
  ylab="tau",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

A contour plot shows where the probability is concentrated. The labels on the contour lines represent how much probabilty is found inside the contour line.

```{r}
density <- density / sum(density, na.rm=TRUE)
density %>%
  as.vector %>%
  sort %>%
  rev -> sorted_density
cumulative_p <- 0
for (p in sorted_density) {
  cumulative_p <- p + cumulative_p
  density[density==p] <- cumulative_p
}
contour(
  theta_grid, 
  tau_grid, 
  density,
  labcex=1,
  levels=c(10, 25, 50, 75, 90)/100,
  xlab="theta",
  ylab="tau")
```

You can see that the hedging hyperprior did its job. The most probable values of $\tau$ are small and the prior mean (`r b0/(a0-1)`) is largely ignored in favor of the mean of the data (`r xbar`).