### 2023-03-12 text/23/normal-hedging-hyperprior.Rmd
Initial graphs for normal-hedging-hyperprior

---
title: "The hedging hyperprior for the normal distribution"
source: "New"
date: "2023-02-28"
categories:
- Blog post
tags:
- Bayesian statistics
output: html_document
---

```{r setup}
suppressMessages(suppressWarnings(library(MASS)))
suppressMessages(suppressWarnings(library(tidyverse)))
knitr::opts_chunk$set(echo=TRUE, fig.width=8, fig.height=2)
```

To do:

+ Rewrite normals using a precision parameter
+ Change the hyperparameter from lambda to delta, kappa, tau, psi, or omega.

I want to write a paper about the hedging hyperprior. I need to work out some simple examples first. Here is an example involving the normal distribution.

<!---more--->

### A simple Baysian analysis with an informative prior.

There are lots of guides on the Internet on using a normal prior for data that is normally distributed with the population standard deviation $\sigma$ known or unknown. I used papers written by [Kevin P. Murphy][mur1] and [Michael I. Jordan][jor1] to get these formulas.

[jor1]: https://people.eecs.berkeley.edu/~jordan/courses/260-spring10/lectures/lecture5.pdf
[mur1]: https://www.cs.ubc.ca/~murphyk/Papers/bayesGauss.pdf

For simplicity, all sums and products are assumed to go from i=1 to n.

Consider a Bayesian analysis of an outcome variable that has a normal distribution. The density of the normal distribution is

$f(X; \mu, \sigma) = \frac{1}{\sigma \sqrt{2 \pi}}e^{-(X-\mu)^2 / 2\sigma^2}$

Assume for simplicity that $\sigma$ is known and equal to $\sigma_F$. A more general case, with $\sigma$ unknown is described on a separate web page.

The mathematical details are a bit easier to follow if you let

$\tau_F = \frac{1}{\sigma_F^2}$

The parameter $\tau$ is called the precision. Note that some define the precision differently as the inverse of the standard deviation rather than the inverse of the variance, The normal density function in terms of $\tau_F$ is

$f(X; \mu, \tau_F) = \sqrt{\frac{\tau_F}{2 \pi}}e^{- \frac{1}{2}\tau_F (X-\mu)^2}$

Simplify this by removing any constants not involving $X$, $\mu$, or $\tau_F$.

$f(X) \propto e^{- \frac{1}{2} \tau_F (X-\mu)^2}$

You have an informative prior on $\mu$ that also has a normal distribution with a prior mean $\mu_0$ and a prior precision $\tau_0$.

$g(\mu; \mu_0, \tau_0) \propto  e^{- \frac{1}{2} \tau_0 (\mu-\mu_0)^2}$

For an informative prior distribution, $\tau_0$ is usually much smaller than $\tau_F$. The ratio $n_0=\tau_F/\tau_0$ is the effective prior sample size and is equivalent to adding $n_0$ pseudo data values of $\mu_0$ to the observed data.

You have collected a sample of n data points, $\mathbf{X} = X_1, X_2, ..., X_n$. This gives you a likelihood of

$L(\mathbf{X} | \mu) \propto e^{- \frac{1}{2} \tau_F\sum(X_i-\mu)^2}$

Rewrite this by incorporating $\bar{X}$ into the squared term

$L(\mathbf{X} | \mu) \propto e^{-\frac{1}{2} \tau_F \sum (X_i-\bar{X}+\bar{X}-\mu)^2}$

When you square the term inside the sum, the cross-product cancels out, leaving you with

$L(\mathbf{X} | \mu) \propto e^{- \frac{1}{2} \tau_F (\sum(X_i-\bar{X})^2} e^{- \frac{n}{2} \tau_F (\bar{X}-\mu)^2}$

$L(\mathbf{X} | \mu) \propto e^{-\frac{n-1}{2}\tau_F S^2} e^{-\frac{n}{2} \tau_F (\mu-\bar{X})^2}$

The posterior distribution is proportional to the product of the prior and the likelihood.

$h(\mu | \mathbf{X}) \propto e^{- \frac{1}{2} \tau_0 (\mu-\mu_0)^2} e^{- \frac{n-1}{2} \tau_F S^2} e^{-\frac{n}{2} \tau_F(\mu-\bar{X})^2)}$

[Note that the following formulas need a lot of work.]

Drop the middle exponential term, which is constant with respect to $\mu$. Set

$\mu_W=\frac{\tau_0 \mu_0 + n \tau_F \bar{X}}{\tau_0 + n \tau_F}$

and rewrite the other two exponentials as

$h(\mu | \mathbf{X}) \propto \tau_0 e^{-\tau_0^2(\mu - \mu_W + \mu_W - mu_0)^2} \tau_F^n e^{-\tau_F(n-1)S^2} e^{-n \tau_F^2(\mu - \mu_W + \mu_W - \bar{X})^2)}$

which simplifies to

$h(\mu | \mathbf{X}) \propto e^{-(n\tau_F+\tau0)^2(\mu-\frac{n \tau_F \bar{X} + \tau_0 \mu_0}{n\tau_F+\tau0})^2}$

which, thanks to the miracle of conjugancy, is a normal distribution. The posterior mean

$\frac{n \tau_F \bar{X} + \tau_0 \mu_0}{n\tau_F+\tau0}$

and the posterior precision is

$n \tau_F + \tau_0$

The posterior mean can be written as a weighted average of the prior mean and the mean of the observed data.

```{r}
x_max <- 10

mu0 <- 8
tau0 <- 4

tauF <- 10

n <- 10
xbar <- 5

mu1 <- (tau0*mu0+n*tauF*xbar)/(tau0+n*tauF)
tau1 <- tau0+n*tauF

x_grid <- seq(0.01, x_max, length=1000)

f <- function(x, mu, tau) {
  sqrt(tau/2*pi)*exp(-0.5*tau*(x-mu)^2)
}

g <- f

prior05 <- qnorm(0.025, mu0, tau0) 
prior95 <- qnorm(0.975, mu0, tau0)

post05 <- qnorm(0.025, mu1, tau1) 
post95 <- qnorm(0.025, mu1, tau1)

```

Let's illustrate this with an informative prior with $\mu_0$ = `r mu0` and $\tau_0$ = `r tau0`.

```{r}
g_grid <- g(x_grid, mu0, tau0)
outside <- x_grid < prior05 | x_grid > prior95
x_color <- 
  ifelse(outside, "white", "gray")
data.frame(x=x_grid, y=g_grid) %>%
  ggplot(aes(x, y)) +
    geom_segment(aes(xend=x, yend=0), color=x_color) +
    geom_line() + xlab(" ") + ylab(" ")
```

This corresponds to a fairly informative prior distribution. This distribution places $\theta$ within `r sprintf("%.2f", prior05)` and `r sprintf("%.2f", prior95)` with 95% probability.

Let's see what happens with this informative prior if you observe `r n` values of x with a mean of `r xbar`. The likelihood is

```{r}
f_grid <- f(xbar, x_grid, tauF)^n
x_color <- "white"
data.frame(x=x_grid, y=f_grid) %>%
  ggplot(aes(x, y)) + 
    geom_line() + xlab(" ") + ylab(" ")
```

Multiply these two together to get the posterior distribution.

```{r}
h_grid <- g(x_grid, mu1, tau1)
outside <- x_grid < post05 | x_grid > post95
x_color <- 
  ifelse(outside, "white", "gray")
data.frame(x=x_grid, y=h_grid) %>%
  ggplot(aes(x, y)) +
    geom_segment(aes(xend=x_grid, yend=0), color=x_color) +
    geom_line() + xlab(" ") + ylab(" ")
```

The posterior distribution is not close to either the prior distribution or the likelihood. It places $\theta$ within `r sprintf("%.2f", post05)` and `r sprintf("%.2f", post95)` with 95% probability. It excludes the prior mean of `r sprintf("%.2f", mu0)` and it excludes `r xbar`, the mean of the data.

### The hedging hyperprior

When the likelihood of your data and your informative prior are in conflict, you may wish to downweight the informative prior. You can do this by including a hedging hyperprior.

Define your prior distribution as an inverse gamma distributions with parameters $1+\lambda (\alpha-1)$ and $\lambda \beta$ where $\lambda$ ranges between 0 and 1. Then place a hyperprior on the parameter $\lambda$. The uniform(0,1) distribution is a reasonable choice for the hyperprior, but others are possible.

```{r}
knitr::opts_chunk$set(echo=TRUE, fig.width=8, fig.height=8)
```

```{r}
lambda_grid <- seq(0, 1, length=60)
mu_grid <- seq(0, x_max, length=60)

gstar <- function(mu, lambda) {g(mu, mu0, lambda*(tau0-tauF)+tauF)}
density <- outer(mu_grid, lambda_grid, gstar)
op <- par(bg = "white") 
z_max <- max(density, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

It is hard to see at this scale, but the prior distribution is higher for small values of $\lambda$ when $\mu$ is very large or very small. This becomes more obvious if you remove the middle of the surface and rescale the Z-axis.

```{r}
density <- outer(mu_grid, lambda_grid, gstar)
density[11:40, ] <- NA
op <- par(bg = "white")
z_max <- max(density, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

The likelihood function is constant across all values of $\tau$, so its surface looks like the following.

```{r}
lambda_grid <- seq(0, 1, length=60)
mu_grid <- seq(0, x_max, length=60)

fstar <- function(mu, lambda) {f(xbar, mu, tau0)^n}
likelihood <- outer(mu_grid, lambda_grid, fstar)
z_max <- max(likelihood, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  likelihood, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

The product of the prior and the likelihood is proportional to the posterior distribution.

```{r}
lambda_grid <- seq(0, 1, length=60)
mu_grid <- seq(0, x_max, length=60)

hstar <- function(mu, lambda) {g(mu, mu0, lambda*(tau0-tauF) + tauF)*f(xbar, mu, tauF)^n}
density <- outer(mu_grid, lambda_grid, hstar)
z_max <- max(density, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

A contour plot shows where the probability is concentrated. The labels on the contour lines represent how much probabilty is found inside the contour line.

```{r}
density <- density / sum(density, na.rm=TRUE)
density %>%
  as.vector %>%
  sort %>%
  rev -> sorted_density
cumulative_p <- 0
for (p in sorted_density) {
  cumulative_p <- p + cumulative_p
  density[density==p] <- cumulative_p
}
contour(
  mu_grid, 
  lambda_grid, 
  density,
  labcex=1,
  levels=c(10, 25, 50, 75, 90)/100,
  xlab="mu",
  ylab="lambda")
```

You can see that the hedging hyperprior did its job. The most probable values of $\lambda$ are small and the prior mean (`r mu0`) is largely ignored in favor of the mean of the data (`r xbar`).