### 2023-03-14 text/23/normal-hedging-hyperprior-part-1.Rmd
Barely started work on normal-hedging-hyperprior-part-2

---
title: "The hedging hyperprior for the normal distribution: unknown mean, known precision"
source: "New"
date: "2023-02-28"
categories:
- Blog post
tags:
- Bayesian statistics
output: html_document
---

```{r setup}
suppressMessages(suppressWarnings(library(MASS)))
suppressMessages(suppressWarnings(library(tidyverse)))
knitr::opts_chunk$set(echo=TRUE, fig.width=8, fig.height=2)
```

I want to write a paper about the hedging hyperprior. I need to work out some simple examples first. Here is an example involving the normal distribution.

<!---more--->

### A simple Baysian analysis with an informative prior.

There are lots of guides on the Internet on using a normal prior for data that is normally distributed with the population standard deviation $\sigma$ known or unknown. I used papers written by [Kevin P. Murphy][mur1] and [Michael I. Jordan][jor1] to get these formulas.

For simplicity, all sums and products are assumed to go from i=1 to n.

Consider a Bayesian analysis of an outcome variable that has a normal distribution. The density of the normal distribution is

For this derivation, it is easier to write the density in terms of

$f(X; \mu, \tau) = \sqrt{\frac{\tau}{2 \pi}}e^{-(X-\mu)^2 / 2\sigma^2}$

where

$\tau=\frac{1}{\sigma*2}$

The parameter $\tau$ is called the precision. Note that some define the precision differently as the inverse of the standard deviation rather than the inverse of the variance,

Assume that $\tau$ is known and equal to $\tau_F$. A more general case, with $\tau$ unknown is described on a separate web page. For simplicity, we will drop any multiplicative constants not involving X or $\mu$

$f(X; \mu) \propto e^{-\frac{1}{2}\tau_F(X-\mu)^2}$

The conjugate prior for a normal distribution is also a normal distribution.

$g(\mu; \mu_0, \tau_0) \propto e^{- \frac{1}{2}\tau_0(\mu-\mu_0)^2}$

where the prior mean is $\mu_0$ and the prior precision is $\tau_0$.

For an informative prior distribution, $\tau_0$ is usually much smaller than $\tau_F$. The ratio $n_0=\tau_F/\tau_0$ is the effective prior sample size and is equivalent to adding $n_0$ pseudo data values of $\mu_0$ to the observed data.

You have collected a sample of n data points, $\mathbf{X} = X_1, X_2, ..., X_n$. This gives you a likelihood of

$L(\mathbf{X} | \mu) \propto e^{- \frac{1}{2} \tau_F\sum(X_i-\mu)^2}$

Rewrite this by incorporating $\bar{X}$ into the squared term

$L(\mathbf{X} | \mu) \propto e^{-\frac{1}{2} \tau_F \sum (X_i-\bar{X}+\bar{X}-\mu)^2}$

When you square the term inside the sum, the cross-product cancels out, leaving you with

$L(\mathbf{X} | \mu) \propto e^{- \frac{1}{2} \tau_F (\sum(X_i-\bar{X})^2} e^{- \frac{n}{2} \tau_F (\bar{X}-\mu)^2}$

$L(\mathbf{X} | \mu) \propto e^{-\frac{n-1}{2}\tau_F S^2} e^{-\frac{n}{2} \tau_F (\mu-\bar{X})^2}$

The posterior distribution is proportional to the product of the prior and the likelihood.

$h(\mu | \mathbf{X}) \propto e^{- \frac{1}{2} \tau_0 (\mu-\mu_0)^2} e^{- \frac{n-1}{2} \tau_F S^2} e^{-\frac{n}{2} \tau_F(\mu-\bar{X})^2)}$

Drop the middle exponential term, which is constant with respect to $\mu$ and use high school algebra ([completing the square][wik1]) to get

$h(\mu | \mathbf{X}) \propto e^{-(n\tau_F+\tau0)(\mu-\frac{n \tau_F \bar{X} + \tau_0 \mu_0}{n\tau_F+\tau0})^2}$

which is a normal distribution. The posterior mean

$\frac{n \tau_F \bar{X} + \tau_0 \mu_0}{n\tau_F+\tau0}$

and the posterior precision is

$n \tau_F + \tau_0$

The posterior mean can be written as a weighted average of the prior mean and the mean of the observed data.

$\frac{n \tau_F}{n\tau_F+\tau_0}\bar{X} + \frac{\tau_0}{n\tau_F+\tau_0}\mu_0$

```{r}
x_max <- 8

mu0 <- 4
tau0 <- 8

tauF <- 0.6
n <- 10
xbar <- 2

mu1 <- (tau0*mu0+n*tauF*xbar)/(tau0+n*tauF)
tau1 <- tau0+n*tauF

x_grid <- seq(0.01, x_max, length=1000)

f <- function(x, mu, tau) {
  sqrt(tau/2*pi)*exp(-0.5*tau*(x-mu)^2)
}

g <- f

prior05 <- qnorm(0.025, mu0, 1/sqrt(tau0)) 
prior95 <- qnorm(0.975, mu0, 1/sqrt(tau0))

post05 <- qnorm(0.025, mu1, 1/sqrt(tau1)) 
post95 <- qnorm(0.975, mu1, 1/sqrt(tau1))

```

Let's illustrate this with an informative prior with $\mu_0$ = `r mu0` and $\tau_0$ = `r tau0`.

```{r}
g_grid <- g(x_grid, mu0, tau0)
outside <- x_grid < prior05 | x_grid > prior95
x_color <- 
  ifelse(outside, "white", "gray")
data.frame(x=x_grid, y=g_grid) %>%
  ggplot(aes(x, y)) +
    geom_segment(aes(xend=x, yend=0), color=x_color) +
    geom_line() + xlab(" ") + ylab(" ")
```

This corresponds to a fairly informative prior distribution. This distribution places $\theta$ within `r sprintf("%.2f", prior05)` and `r sprintf("%.2f", prior95)` with 95% probability.

Let's see what happens with this informative prior if you observe `r n` values of x with a mean of `r xbar`. The likelihood is

```{r}
f_grid <- f(xbar, x_grid, tauF)^n
x_color <- "white"
data.frame(x=x_grid, y=f_grid) %>%
  ggplot(aes(x, y)) + 
    geom_line() + xlab(" ") + ylab(" ")
```

Multiply these two together to get the posterior distribution.

```{r}
h_grid <- g(x_grid, mu1, tau1)
outside <- x_grid < post05 | x_grid > post95
x_color <- 
  ifelse(outside, "white", "gray")
data.frame(x=x_grid, y=h_grid) %>%
  ggplot(aes(x, y)) +
    geom_segment(aes(xend=x_grid, yend=0), color=x_color) +
    geom_line() + xlab(" ") + ylab(" ")
```

The posterior distribution is not close to either the prior distribution or the likelihood. It places $\theta$ within `r sprintf("%.2f", post05)` and `r sprintf("%.2f", post95)` with 95% probability. It excludes the prior mean of `r mu0` and it excludes `r xbar`, the mean of the data.

### The hedging hyperprior

When the likelihood of your data and your informative prior are in conflict, you may wish to downweight the informative prior. You can do this by including a hedging hyperprior.

Define your prior distribution as a normal distributions with prior mean $\mu0$ and prior precision

$\tau_F+\lambda (\tau_0-\tau_F)$

where $\lambda$ ranges between 0 and 1 and $\tau_F$ is less than $\tau_0$. When $\lambda=0$, the prior precision reduces to $\tau_F$ which is equivalent to a prior sample size of $n_0=1$. When $lambda=1$, the prior precision is $\tau_0$ which represents the full strength of the informative priori distribution.

Place a hyperprior on the parameter $\lambda$. The uniform(0,1) distribution is a reasonable choice for the hyperprior, but others are possible.

```{r}
knitr::opts_chunk$set(echo=TRUE, fig.width=8, fig.height=8)
```

```{r}
lambda_grid <- seq(0, 1, length=60)
mu_grid <- seq(0, x_max, length=60)

gstar <- function(mu, lambda) {g(mu, mu0, lambda*(tau0-tauF)+tauF)}
density <- outer(mu_grid, lambda_grid, gstar)
op <- par(bg = "white") 
z_max <- max(density, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

This is the joint prior distribution of $\mu_0$ and $\lambda$.

It is hard to see at this scale, but the prior distribution is higher for small values of $\lambda$ when $\mu$ is very large or very small. This becomes more obvious if you remove the middle of the surface and rescale the Z-axis.

```{r}
density <- outer(mu_grid, lambda_grid, gstar)
density[16:44, ] <- NA
op <- par(bg = "white")
z_max <- max(density, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

The likelihood function is constant across all values of $\tau$, so its surface looks like the following.

```{r}
lambda_grid <- seq(0, 1, length=60)
mu_grid <- seq(0, x_max, length=60)

fstar <- function(mu, lambda) {f(xbar, mu, tau0)^n}
likelihood <- outer(mu_grid, lambda_grid, fstar)
z_max <- max(likelihood, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  likelihood, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

The product of the prior and the likelihood is proportional to the posterior distribution.

```{r}
lambda_grid <- seq(0, 1, length=60)
mu_grid <- seq(0, x_max, length=60)

hstar <- function(mu, lambda) {g(mu, mu0, lambda*(tau0-tauF) + tauF)*f(xbar, mu, tauF)^n}
density <- outer(mu_grid, lambda_grid, hstar)
z_max <- max(density, na.rm=TRUE)
persp(
  mu_grid, 
  lambda_grid, 
  density, 
  zlim=c(0, 1.1*z_max),
  xlab="mu",
  ylab="lambda",
  theta = 135, 
  phi = 30, 
  expand=0.5,
  ticktype="detailed")
```

A contour plot shows where the probability is concentrated. The labels on the contour lines represent how much probabilty is found inside the contour line.

```{r}
density <- density / sum(density, na.rm=TRUE)
density %>%
  as.vector %>%
  sort %>%
  rev -> sorted_density
cumulative_p <- 0
for (p in sorted_density) {
  cumulative_p <- p + cumulative_p
  density[density==p] <- cumulative_p
}
contour(
  mu_grid, 
  lambda_grid, 
  density,
  labcex=1,
  levels=c(50, 75, 90)/100,
  xlab="mu",
  ylab="lambda")
```

You can see that the hedging hyperprior did its job. The most probable values of $\lambda$ are small and the prior mean (`r mu0`) is largely ignored in favor of the mean of the data (`r xbar`).

[jor1]: https://people.eecs.berkeley.edu/~jordan/courses/260-spring10/lectures/lecture5.pdf
[mur1]: https://www.cs.ubc.ca/~murphyk/Papers/bayesGauss.pdf
[wik1]: https://en.wikipedia.org/wiki/Completing_the_square