### 2023-05-24 src/check_web.Rmd
Added links to blog pages from 2014

---
title: "Check web files for links to new pages"
author: "Steve Simon"
date: "Created 2022-05-26"
output: html_document
---

### Setup

```{r setup, echo=TRUE}
knitr::opts_chunk$set(echo=TRUE)
suppressMessages(suppressWarnings(library(glue)))
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(lubridate)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(tidyverse)))
```

### Read html files

```{r read_old}
path_prefix <- "c:/Users"
path_suffix <- "dropbox/professional/fun stuff/update-website"

list.files(path_prefix) %>%
  str_subset("(Simons)|(steve)") -> me

full_path <- glue("{path_prefix}/{me}/{path_suffix}")

subdir_list=c("99", paste0("0", 0:9), 10:22)
html_files <- NULL
for (yr in rev(subdir_list)) {
  file_path <- paste0(full_path, "/", yr)
  yr_files <- list.files(file_path, pattern="html")
  if (length(yr_files)==0) next
  for (fn in yr_files) {
    read_lines(paste0(file_path, "/", fn)) %>%
      str_subset('<p>This page (is moving)|(has moved)') %>%
      str_remove('.*<a href="http://new.pmean.com/') %>%
      str_remove('/">.*') %>%
      str_remove('<p>This page is being ') %>%
      str_remove('.</p>') %>%
      str_trim -> link_file
    if (length(link_file)==0) link_file <- "NOTHING"
    cat(glue("\n\n{yr}/{fn}"))
    cat(glue(" links to {link_file}."))
  }
}
```