### 2023-07-24 src/year-end-summary.Rmd
Updated year-end-summary

---
title: "Create year end summaries"
author: "Steve Simon"
date: "Created 2022-05-26"
output: html_document
---

This program produces Markdown files summarizing every webpage written in a given year

```{r Preliminaries}
suppressMessages(suppressWarnings(library(glue)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(readr)))
suppressMessages(suppressWarnings(library(stringr)))

# pad single digits with leading zero
zpad <- function(x) {
  ifelse(nchar(x)==1, paste0("0", x), x)
}

update <- TRUE
print("Check 1")
```

```{r}
# Create yaml header
create_yaml <- function(yyyy) {
  glue(
    '---
     title: "Pages written in {yyyy}"
     source: new
     date: {yyyy}-12-31
     categories:
     - Blog post
     tags:
     - Website details
     output: html_document
     ---')
}
```

```{r}
# Create links to other years
create_links <- function(yyyy_range) {
  yy_range <- str_sub(yyyy_range, 3)
  glue("[{yyyy_range}](http://new.pmean.com/{yyyy_range}/)") %>%
    paste(collapse=" | ")
}
```

```{r}
# Check for problems

# Check source
check_source <- function(file_text) {
  message <- NULL
  file_text %>%
    str_subset("^source\\: ") %>%
    str_remove(fixed("source: ")) %>%
    str_remove_all('"')  %>%
    str_remove_all("'") -> file_source
  if (length(file_source)==0) {
    message <- glue("{yy}/{fn} has no source\n")
  }
  if (length(file_source)>1) {
    message <- glue("{yy}/{fn} has multiple sources\n")
  }
  return(message)
}
```

```{r}
# Check date
check_date <- function(file_text) {
  message <- NULL
  file_text %>%
    str_subset("^date\\: ") %>%
    str_remove(fixed("date: ")) %>%
    str_remove_all('"') %>%
    str_remove_all("'")  -> message
  if (length(message)==0) {message <- glue("{yy}/{fn} has no date")}
  if (length(message)>1) {message <- glue("{yy}/{fn} has multiple dates")}
  return(message)
}

# Check title
```

```{r}
# Create links for each file
create_file_list <- function(f, date_list) {
  f %>%
    str_replace(fixed(".Rmd"), "") %>%
    str_replace(fixed(".md"), "") -> short_name
  o <- order(date_list, f)
  glue("+ {date_list[o]} [{short_name[o]}](http://new.pmean.com/{short_name[o]}/)  ") %>%
    paste(collapse="  \n")
}
```

```{r}
# Put it all together
yyyy_list <- c(1999:2023)
n_problems <- 0
n_files <- 0
for (yyyy in yyyy_list) {
  yy <- str_sub(yyyy, 3, 4)
  output_file <- glue("../text/{yy}/{yyyy}.md")
  glue("../text/{yy}") %>%
    list.files %>%
    str_subset("md$") -> f
  problem_files <- NULL
  date_list <- NULL
  for (fn in f) {
    file_text <- read_lines(glue("../text/{yy}/{fn}"))
    problem_files %<>% append(check_source(file_text))
    date_list %<>% append(check_date(file_text))
  }

  # Update if no problems found
  date_list %>% 
    str_subset(" for ") %>%
    append(problem_files) -> problem_files

  n_year <- length(problem_files)  
  n_problems <- n_problems + n_year
  n_files <- n_files + length(f)
  if (n_year==0) {
    problem_files <- "All files passed date and source quality checks"
  }

  if (n_year > 0) {print(problem_files)}

  cat("\n")
  cat(glue("{yyyy} has a total of {length(f)} files."))
  cat("\n\n")

  if (update==FALSE) {next}

  create_yaml(yyyy) %>% write_lines(output_file)
  write_lines(" ", output_file, append=TRUE)

  create_links(yyyy_list) %>% write_lines(output_file, append=TRUE)
  write_lines(" ", output_file, append=TRUE)

  glue("New website ({length(f)} files)") %>% write_lines(output_file, append=TRUE)
  write_lines(" ", output_file, append=TRUE)

  create_file_list(f, date_list) %>% write_lines(output_file, append=TRUE)
  write_lines(" ", output_file, append=TRUE)

  problem_files %>% write_lines(output_file, append=TRUE)

  glue("{yyyy}.md created, {length(f)} files, {n_year} problems detected") %>% print
}

print(glue("All years: {n_problems} problems identified in {n_files} files"))
```