### 2024-05-09 text/24/survival-spline.Rmd
Added survival-spline

---
title: Illustrating splines using the Worcester Heart Attack Study
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Survival analysis
- R programming
output: html_document
---

Splines provide a useful way to model relationships that are more complex than a simple linear relationship. They work in a variety of regression models. Here is an illustration of how to use a spline in a Cox regression model with data from the Worcester Heart Attack Study.

<!---more--->

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
f0 <- "https://raw.githubusercontent.com/pmean/datasets/master/"
f1 <- "whas100-data-dictionary.yaml"
dd <- read_yaml(paste0(f0, f1))
```

Here is a brief description of the whas100 dataset, taken from the [data dictionary][sim3] on my github site.

`r dd$description`

[sim3]: https://raw.githubusercontent.com/pmean/datasets/master/whas100-data-dictionary.yaml

Here are the first few rows of data and the last few rows of data. Row 101 needs to be removed.

```{r}
f2 <- "whas100.dat"
raw_data <- read.fwf(
  file=paste0(f0, f2),
  widths=c(7, 12, 12, 5, 9, 5, 7, 4, 8))
wa <- raw_data
names(wa) <- c(
  "id",
  "admitdate",
  "foldate",
  "los",
  "lenfol",
  "fstat",
  "age",
  "gender",
  "bmi")
head(wa)
tail(wa)
wa <- wa[-101, ]
wa$lenfol <- wa$lenfol/365.25
```

Here are a few descriptive statistics

```{r}
summary(wa$id)
summary(wa$los)
summary(wa$lenfol)
count(wa, fstat)
summary(wa$age)
count(wa, gender)
summary(wa$bmi)
```

Here is an overall survival curve.

```{r}
plot(Surv(wa$lenfol, wa$fstat))
```

```{r}
linear_fit <- coxph(
  Surv(wa$lenfol, wa$fstat)~age,
  data=wa)
linear_fit
```

```{r}
spline_fit <- coxph(
  Surv(wa$lenfol, wa$fstat)~rcs(age),
  data=wa)
spline_fit
```

```{r}
age_sequence <- data.frame(
  age = seq(min(wa$age), max(wa$age), length=100))
age_sequence$prediction <- 
  predict(spline_fit, newdata=age_sequence, type="lp")
ggplot(data=age_sequence, aes(age, prediction)) +
  geom_line()
```