### 2024-05-09 text/pending/survival-lasso.Rmd
Started work on survival-lasso

---
title: Illustrating the lasso model using the Worcester Heart Attack Study
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Survival analysis
- R programming
output: html_document
---

The lasso model allows you to effectively identify important features in a large regression model. Here is an illustration of how to use a the lasso model with survival data from the Worcester Heart Attack Study.

<!---more--->

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(glmnet)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
f0 <- "https://raw.githubusercontent.com/pmean/datasets/master/"
f1 <- "whas500-data-dictionary.yaml"
dd <- read_yaml(paste0(f0, f1))
```

Here is a brief description of the whas100 dataset, taken from the [data dictionary][sim3] on my github site.

`r dd$description`

[sim3]: https://raw.githubusercontent.com/pmean/datasets/master/whas100-data-dictionary.yaml

Here are the first few rows of data and the last few rows of data. Row 101 needs to be removed.

```{r}
vnames <- names(dd$vars)
f2 <- "whas500.dat"
w0 <- read_table(
  file=paste0(f0, f2), 
  col_names=c(vnames),
  col_types="nnnnnnnnnnnnnnncccnnnn")
head(w0)
```

Here are a few descriptive statistics

```{r}
dd$vars[["id"]]
summary(w0$id)

dd$vars[["age"]]
summary(w0$age)

dd$vars[["gender"]]
count(w0, gender)

dd$vars[["sysbp"]]
summary(w0$sysbp)

dd$vars[["diasbp"]]
summary(w0$diasbp)

dd$vars[["bmi"]]
summary(w0$bmi)

dd$vars[["cvd"]]
count(w0, cvd)

dd$vars[["afb"]]
count(w0, afb)

dd$vars[["sho"]]
count(w0, sho)

dd$vars[["chf"]]
count(w0, chf)

dd$vars[["av3"]]
count(w0, av3)

dd$vars[["miord"]]
count(w0, miord)

dd$vars[["mitype"]]
count(w0, mitype)

dd$vars[["year"]]
count(w0, year)

dd$vars[["los"]]
summary(w0$los)

dd$vars[["dstat"]]
count(w0, dstat)

dd$vars[["lenfol"]]
summary(w0$lenfol)

dd$vars[["fstat"]]
count(w0, fstat)
```

Here is an overall survival curve.

```{r}
plot(Surv(w0$lenfol, w0$fstat))
```

You can find the lasso regression model in the glmnet package. One important difference with glmnet is that it does not use a formula to define your model. Instead you provide matrices for your dependent variable and your independent variables.