### 2024-05-11 text/pending/survival-lasso.Rmd
First full draft of survival-lasso

---
title: Illustrating the lasso model using the Worcester Heart Attack Study
source: "New"
author: Steve Simon
date: 2024-05-07
categories:
- Blog post
tags:
- Survival analysis
- R programming
output: html_document
---

The lasso model allows you to effectively identify important features in a large regression model. Here is an illustration of how to use a the lasso model with survival data from the Worcester Heart Attack Study.

I've followed the [glmnet vignette for Cox regression][tay1] closely. You might find that explanation to be clearer than what I present below.

[tay1]: https://glmnet.stanford.edu/articles/Coxnet.html

<!---more--->

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(broom)
library(glmnet)
library(glue)
library(pspline)
library(readr)
library(rms)
library(tidyverse)
library(yaml)
g0 <- "https://raw.githubusercontent.com/pmean/"
r0 <- "datasets/master/"
s1 <- "yaml-functions.R"
f1 <- "whas500.yaml"

source(paste0(g0, r0, s1))
dd <- read_yaml(paste0(g0, r0, f1))
vnames <- extract_names(dd)
vlabels <- extract_labels(dd)
vscales <- extract_scales(dd)
```

Here is a brief description of the whas500 dataset, taken from the [data dictionary][sim3] on my github site.

`r dd$description`

[sim3]: https://raw.githubusercontent.com/pmean/datasets/master/whas100-data-dictionary.yaml

Here are the first few rows of data and the last few rows of data. Row 101 needs to be removed.

```{r}
vnames <- names(dd$vars)
f2 <- "whas500.dat"
w0 <- read_table(
  file=paste0(g0, r0, f2), 
  col_names=c(vnames),
  col_types="nnnnnnnnnnnnnnncccnnnn")
head(w0)
```

Here are a few descriptive statistics

```{r}
categorical <- 
  (vscales=="binary") |
  (vscales=="nominal") |
  (vscales=="ordinal")
for (i_vars in which(categorical)) {
  cat(glue("{vnames[i_vars]}: {vlabels[i_vars]}"), "\n")
  w0 %>%
    count(.data[[vnames[i_vars]]]) %>%
    data.frame %>%
    print
  cat("\n")
}
```

```{r}
continuous <- 
  (vscales=="interval") |
  (vscales=="ratio")
for (i_vars in which(continuous)) {
  cat(glue("{vnames[i_vars]}: {vlabels[i_vars]}"), "\n")
  print(summary(w0[vnames[i_vars]]))
  cat("\n")
}
```

Here is an overall survival curve.

```{r}
plot(Surv(w0$lenfol, w0$fstat))
```

You can find the lasso regression model in the glmnet package. One important difference with glmnet is that it does not use a formula to define your model. Instead you provide matrices for your dependent variable and your independent variables. In a Cox regression model, the dependent matrix has a column for time and a second column with an indicator variable (1 means the event occurred and 0 means censoring).

```{r}
ymat <- cbind(w0$lenfol, w0$fstat)
dimnames(ymat)[2] <- list(c("time", "status"))
xmat <- as.matrix(w0[ , 8:14])
dimnames(xmat)[2] <- list(vnames[8:14])
lasso_fit <- glmnet(xmat, ymat, family="cox")
plot(lasso_fit, label=TRUE)
plot(lasso_fit, xvar="norm", label=TRUE)
plot(lasso_fit, xvar="lambda", label=TRUE)
plot(lasso_fit, xvar="dev", label=TRUE)
```

```{r}
cv_fit <- cv.glmnet(xmat, ymat, family = "cox", type.measure="mae")
plot(cv_fit)
plot(lasso_fit, xvar="lambda")
abline(v=log(cv_fit$lambda.1se))
cv_fit$lambda.1se
coef(lasso_fit, cv_fit$lambda.1se)
```

```{r}
cv_fit$lambda.min
plot(lasso_fit, xvar="lambda")
abline(v=log(cv_fit$lambda.min))
coef(lasso_fit, cv_fit$lambda.min)
```