### 2024-06-21 src/count_completed.Rmd
Update 2006 files

---
title: "Count pages completed in each year"
author: "Steve Simon"
date: "Created 2024-06-16"
output: html_document
---

### Setup

```{r setup, echo=TRUE}
knitr::opts_chunk$set(echo=TRUE)
suppressMessages(suppressWarnings(library(glue)))
suppressMessages(suppressWarnings(library(knitr)))
suppressMessages(suppressWarnings(library(lubridate)))
suppressMessages(suppressWarnings(library(tidyverse)))
```

### Read md files

```{r read_md}
md_path <- "../text/"
subdir_list=c("99", paste0("0", 0:9), 10:24)
md_files <- NULL
for (yr in rev(subdir_list)) {
  file_path <- paste0(md_path, yr)
  file_path %>%
    list.files(pattern="md$") %>%
    data.frame(yr=yr, fn=.) %>%
    bind_rows(md_files) -> md_files
}
count(md_files, yr)
```

### List duplicates

```{r list-duplicates}
md_files %>%
  group_by(fn) %>%
  mutate(n=n()) %>%
  filter(n>1) %>%
  select(-n) %>%
  arrange(fn, yr) %>%
  data.frame
```

### Count page_update fields

```{r md_links}
c_text <- "page_update: "
md_files$page_update_line <- "999"
for (i_row in 1:nrow(md_files)) {
  full_file_name <- glue("{md_path}/{md_files$yr[i_row]}/{md_files$fn[i_row]}")
  read_lines(full_file_name) %>%
    str_subset(fixed(c_text)) %>%
    str_replace(c_text, "") -> page_update
  n <- length(page_update)
  if (n > 1) {
    cat(
      glue(
        "\n\n{n} duplicates found in {full_file_name}.\n\n"))}
  page_update %>%
    paste0("", collapse=", ") -> md_files$page_update_line[i_row]
}
count(md_files, page_update_line)
count(md_files, yr, page_update_line)
```

```{r save_everything}
save(md_files, file="../data/page_update_check.RData")
```