### 2024-06-27 src/audit-cross-links.Rmd
Added nice-layout

---
title: "Check links on new pages"
author: "Steve Simon"
date: "Created 2024-06-25"
output: html_document
---

This program looks for key information in the new webpages

```{r setup}
suppressMessages(suppressWarnings(library(glue)))
suppressMessages(suppressWarnings(library(magrittr)))
suppressMessages(suppressWarnings(library(readr)))
suppressMessages(suppressWarnings(library(stringr)))
suppressMessages(suppressWarnings(library(tidyverse)))

load(file="../data/check_new.RData")
n_new <- nrow(md)
load(file="../data/check_old.RData")
n_old <- nrow(html)
size_limit <- 99

inc <- function(x) {
  n <- length(x)
  m <- c(FALSE, x[-1]==x[-n])
  x[m] <- " "
  x
}
test1 <- rep(letters[1:4], 1:4)
test2 <- inc(test1)
data.frame(test1, test2)
```

### pages with no links from new site to old or blog site

```{r no-link-www}
md %>% 
  filter(miss==1) %>%
  select(yr, fn, stat) -> pip1 
print(nrow(pip1))
glue("{nrow(pip1)} / {n_new} = {round(100*nrow(pip1)/n_new)}%")
pip1 %>%
  mutate(yr=inc(yr)) %>%
  head(size_limit)
```

### pages with bad links from new site to old site

```{r bad-link-www}
html %>%
  rename(base=fn) %>%
  rename(subd=yr) %>%
  select(base, subd, stub) -> html1
md %>% 
  filter(miss==0) %>%
  filter(site=="www") %>%
  left_join(html1, by=c("subd", "base")) %>%
  select(yr, fn, stat, subd, base, stub) -> pip1
pip1 %>%
  filter(is.na(stub)) %>%
    select(-stub) -> pip2 
glue("{nrow(pip2)} / {n_new} = {round(100*nrow(pip2)/n_new)}%")
pip2 %>%
  mutate(yr=inc(yr)) %>%
  head(size_limit)
```

### pages with good links from new site to old site

```{r good-link-www}
pip1 %>%
  filter(!is.na(stub))  -> pip3 
glue("{nrow(pip3)} / {n_new} = {round(100*nrow(pip3)/n_new)}%")
pip3 %>%
  mutate(yr=inc(yr)) %>%
  head(size_limit)
```

### pages with no links from old site to new site

```{r no-link-new}
html %>% 
  filter(miss==1) %>%
  select(yr, fn, stub) -> pip1 
print(nrow(pip1))
pip1 %>%
  mutate(yr=inc(yr)) %>%
  head(size_limit)
```

### pages with bad links from old site to new site

```{r bad-link-new}
md %>%
  select(-link) %>%
  rename(link=fn) %>%
  select(link, stat) -> md1

html %>%
  filter(miss==0) %>%
  left_join(md1, by="link") %>%
  select(yr, fn, link, stat) -> pip1
pip1 %>%
  filter(is.na(stat)) %>%
  select(-stat)  -> pip4 
print(nrow(pip4))
pip4 %>%
  mutate(yr=inc(yr)) %>%
  head(size_limit)
```

### pages with good links from old site to new site

```{r good-link-new}
pip1 %>%
  filter(!is.na(stat)) -> pip5 
print(nrow(pip5))
pip5 %>%
  mutate(yr=inc(yr)) %>%
  head(size_limit)
```