### 2024-07-06 text/22/testing-d3.Rmd

---
title: "Testing d3"
author: "Steve Simon"
date: "2022-06-01"
source: new
categories:
- Blog post
tags:
- Graphical display
- Rmarkdown graphs
output: html_document
page_update: complete
---

I am interested in developing some visualization applications using d3. You can run d3 from within R Markdown.

This is an example that I borrowed from [Section 15.5][xie1] of [R Markdown Cookbook][xie2] by Yihui Xie, Christophe Dervieus, and Emily Riederer.

First, load the package **r2d3** to set up the `d3` engine
for **knitr** automatically:

```{r testing-d3-01}
library(r2d3)
```

Now we can generate data in R, pass it to D3, and draw
the chart:

```{d3 testing-d3-02, data=runif(30), options=list(color='steelblue')}
svg.selectAll('rect')
  .data(data)
  .enter()
    .append('rect')
      .attr('width', function(d) { return d * 672; })
      .attr('height', '10px')
      .attr('y', function(d, i) { return i * 16; })
      .attr('fill', options.color);
```

[xie1]: https://bookdown.org/yihui/rmarkdown-cookbook/d3.html
[xie2]: https://bookdown.org/yihui/rmarkdown-cookbook/