### 2024-07-06 text/21/venn-diagram.Rmd

---
title: "Test diagrams"
author: "Steve Simon"
source: new
date: "2021-10-30"
categories:
- Blog post
tags:
- Graphical display
- Rmarkdown graphs
output: html_document
page_update: complete
---

```{r venn-diagrams-01, echo=FALSE}
suppressMessages(suppressWarnings(library(dplyr)))
suppressMessages(suppressWarnings(library(ggplot2)))
suppressMessages(suppressWarnings(library(magrittr)))
```

```{r venn-diagrams-02, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.6
v <- 0.85
draw_circles <- function() {
  dx <- 1
  dy <- 0.866
  i <- seq(0, 2*pi, length=100)
  circle <- data.frame(x=sin(i), y=cos(i))
  circle %>%
    mutate(g=1) %>%
    mutate(x=x-o*dx) %>%
    mutate(y=y+o*dy) -> circle1
  circle %>%
    mutate(g=2) %>%
    mutate(x=x+o*dx) %>%
    mutate(y=y+o*dy) -> circle2
  circle %>%
    mutate(g=3) %>%
    mutate(y=y-o*dy) -> circle3

  circle1 %>%
    bind_rows(circle2) %>%
    bind_rows(circle3) %>%
    ggplot(aes(x, y, group=g)) + 
      geom_path(color=rep(c("red", "darkgreen", "blue"), each=100), size=3) +
      geom_text(x=v*dx, y=v*dy, size=5, label="Math,\nStat", color="darkgreen") +
      geom_text(x=-v*dx, y=v*dy, size=5, label="Computers,\nCoding", color="red") +
      geom_text(x=0, y=-v*dy, size=5, label="Science,\nMedicine", color="blue") +
      expand_limits(x=c(-(dx+1),(dx+1))) +
      expand_limits(y=c(-(dx+1),(dx+1))) +
      theme_void()
}
o <- 0.6
v <- 0.85
draw_circles()
```

```{r venn-diagrams-03, fig.width=5, fig.height=5, echo=FALSE}
o <- 0.88
draw_circles()
```