### 2025-02-09 text/25/pretty-functions.Rmd
Finalized pretty-functions

---
title: "Functions to make plain R output look nicer"
author: "Steve Simon"
source: new
date: "2025-02-09"
categories:
- Blog post
tags:
- R programming
output: html_document
page_update: complete
---

Most of the statistical summary functions in R produce plain vanilla output. That's actually good because it allows you to customize things the way you like. I have been finding that I use pretty close to the same customizations over and over again, so I thought I should standardize them in a few simple R functions.

<!---more--->

```{r setup, message=FALSE, warning=FALSE, echo=FALSE}
library(broom)
library(glue)
library(palmerpenguins)
# data(package = 'palmerpenguins')
library(tidyverse)
```

## pretty_p

The pretty_p function represents my efforts to avoid displaying small p-values using scientific notation. Some researchers get confused when they see something like 7.93e- 2. Even for those researchers who are used to scientific notation, it just makes things hard to read. This function replaces very small p-values with "p < 0.001" and rounds larger p-values to two significant figures.

```{r}
#| label: function-1

pretty_p <- function(p_table) {
  # p_table is the output from the tidy function in the broom package
  p_table |>
    mutate(p.value =
      case_when(
        p.value <  0.001 ~ "p < 0.001",
        p.value >= 0.001 ~ glue("p = {signif(p.value, 2)}")))
}
```

#### Comment on the code

The [mutate][ref01] function modifies an existing variable or creates a new variable. The [case_when][ref02] function uses logic statements to assign values. Both are part of the [dplyr][ref03]/[tidyverse][ref04] libraries. The [signif][ref05] function, part of the base R package rounds a value to a specified number of significant digits. The [glue][ref06] function, part of the [glue][ref07] package, will insert variables inside of a string by surrounding the variable by curly brackets.

[ref01]: https://dplyr.tidyverse.org/reference/mutate.html
[ref02]: https://dplyr.tidyverse.org/reference/case_when.html
[ref03]: https://dplyr.tidyverse.org/
[ref04]: https://www.tidyverse.org/
[ref05]: https://stat.ethz.ch/R-manual/R-devel/library/base/html/Round.html
[ref06]: https://glue.tidyverse.org/
[ref07]: https://glue.tidyverse.org/articles/glue.html

#### Example

Here is an example of how you might use the pretty_p function. I am using the [Palmer Penguins][ref08] dataset. Here is the output from tidy without pretty_p.

[ref08]: https://allisonhorst.github.io/palmerpenguins/

```{r}
lm(flipper_length_mm ~ species + island, data=penguins) |>
  tidy()
```

Here is the output from tidy with pretty_p.

```{r}
lm(flipper_length_mm ~ species + island, data=penguins) |>
  tidy() |>
  pretty_p()
```

## pretty_n

I also like to replace counts with percentages, but show the numerator and denominator in parentheses afterwards. This function will take results from the count function (or from any dataset with a variable named "n"), compute a total and percentage, and then display the results in a nice format.

```{r}
pretty_n <- function(n_table) {
  # n_table is output from the count function
  n_table |>
  	mutate(total=sum(n)) |>
  	mutate(pct=round(100*n/total)) |>
	  mutate(pct=glue("{pct}% ({n}/{total})")) |>
	  select(-n, -total)
}
```

#### Comment on the code

The mutate and glue functions are described above. The round function is described in the same link as the signif function. The [select][ref09] function will keep or drop columns of data.

[ref09]: https://dplyr.tidyverse.org/reference/select.html

#### Example

Here is an example of how you might use the pretty_n function. Here is the output from count without pretty_n.

```{r}
penguins |>
  count(species)
```

Here is the output from count with pretty_n.

```{r}
penguins |>
  count(species) |>
  pretty_n()
```

## pretty_mean

For continuous variables, I usually like to compute a few descriptive statistics: the mean, standard deviation, minimum, maximum, and the number of missing values. This is easy enough to do but the code is long and tedious. The pretty_mean function is just as long and tedious (if not more so), but once you enter it, you can save a lot of time starting with the second variable that you need to summarize. I actually used to like the [summary][ref10] function, but it does not include a standard deviation and it does not work well with the [group_by][ref11] function in dplyr.

[ref10]: https://stat.ethz.ch/R-manual/R-devel/library/base/html/summary.html
[ref11]: https://dplyr.tidyverse.org/reference/group_by.html

```{r}
pretty_mean <- function(d, v) {
  # d is a data frame or tibble, v is a variable in d.
  d |>
    summarize(n_missing=sum(is.na({{v}}))) -> d1
  d |>
    filter(!is.na({{v}})) |>
    summarize(
      across({{v}}, 
        list(
          mean=mean, 
          sd=sd,
          min=min,
          max=max))) |>
    bind_cols(d1)
}
```

#### Comment on the code

Using tidyverse functions [inside a loop or function][ref12] is tricky. With an argument that represents a variable name, you either enclose the argument in double curly braces ("embrace" is the term they use) or you specify the argument inside of .data[[]].

The [across][ref13] function, part of the dplyr/tidyverse libraries, allows you to select specific variable(s) for further operations within dplyr functions like summarize. I have had some difficulty with this function, but it seems to work well here.

The [bind_cols][ref14] function, part of the dplyr library, combines two data frames or tibbles side by side.

[ref12]: https://dplyr.tidyverse.org/articles/programming.html
[ref13]: https://dplyr.tidyverse.org/reference/across.html
[ref14]: https://dplyr.tidyverse.org/reference/bind_cols.html

#### Example

Here is the code you would write without the pretty_mean function.

```{r}
penguins |>
  summarize(
    body_mass_g_mean=mean(body_mass_g, na.rm=TRUE),
    body_mass_g_sd=sd(body_mass_g, na.rm=TRUE),
    body_mass_g_min=min(body_mass_g, na.rm=TRUE),
    body_mass_g_max=max(body_mass_g, na.rm=TRUE),
    n_missing=sum(is.na(body_mass_g)))
```

Here is the code with the pretty_mean function.

```{r}
penguins |>
  pretty_mean(body_mass_g)
```

Writing functions like these only make sense if you find yourself doing the same sort of thing three or more times in one program. There is a nice side effect of writing the function. If you decide that you want a different appearance or you want to round some of your results or you want to feed the data into a [nicely formatted table][ref15], it only takes a small modification in one spot.

[ref15]: https://cran.r-project.org/web/packages/insight/vignettes/display.html