### 2025-03-24 src/copy-files.qmd
Updated copy-files to also read 2025 pages

---
title: "Copy files from build-website to qblog"
author: "Steve Simon"
format:
  html:
    embed-resources: true
---

This program was written by Steve Simon on 2024-11-09 and is placed in the public domain.

## Preliminaries

```{r}
#| label: setup
#| message: false
#| warning: false

library(glue)
library(tidyverse)
```

## Set old and new locations

```{r}
#| label: locations
old_path <- "../text"
old_directories <- c("99", paste0("0", (0:9)), 10:25)
new_path <- "../../qblog/posts"
file_pattern <- "md$"
```

## Find files function

```{r}
#| label: find-files
files <- NULL
subdirectories <- NULL
for (dir in old_directories) {
  glue("{old_path}/{dir}") |>
    list.files() |>
    str_subset(file_pattern) -> to_be_copied

  n_to_be_copied <- length(to_be_copied)
  subdirectories <- append(subdirectories, rep(dir, n_to_be_copied))
  files <- append(files, to_be_copied)

  message <- glue(
    "Searching through {n_to_be_copied} files in {old_path}/{dir}")
  cat(message)
  cat("\n")
}

files |>
  str_remove("\\..*") -> roots

files |> 
  str_remove("^.*\\.") -> extensions
```

```{r}
#| label: create-file-names
new_paths <- glue("{new_path}/{roots}")
new_files <- glue("{new_paths}/index.{extensions}")
old_files <- glue("{old_path}/{subdirectories}/{roots}.{extensions}")
```

```{r}
#| label: move
n_files <- length(files)
for (i in 1:n_files) {
  if (!dir.exists(new_paths[i])) {dir.create(new_paths[i])}
  test1 <- file.copy(from=old_files[i], to=new_files[i], overwrite=TRUE)
  cat(glue("{test1}: {old_files[i]} copied to {new_files[i]}"))
  cat("\n")
  cat(list.files(new_paths[i]))
  cat("\n")
}
cat(str_glue("Total files copied: {n_files}"))
```