### 2025-03-24 src/rmd-check.qmd

---
title: "Check that all Rmd files render properly"
author: "Steve Simon"
format:
  html:
    embed-resources: true
---

This program was written by Steve Simon on 2024-11-09 and is placed in the public domain.

## Preliminaries

```{r}
#| label: setup
#| message: false
#| warning: false

library(glue)
library(quarto)
library(tidyverse)
```

## Set old and new locations

```{r}
#| label: locations

old_path <- "../text"
old_directories <- c(25:10, paste0("0", (9:0)), "99")
new_path <- "../../qblog/posts"
file_pattern <- "Rmd$"
```

## Find files

```{r}
#| label: find-files

rmd_files <- NULL
for (dir in old_directories) {
  glue("{old_path}/{dir}") |>
    list.files() |>
    str_subset(file_pattern) -> to_be_tested

  n_to_be_tested <- length(to_be_tested)
  message <- glue(
    "Examining {n_to_be_tested} files in {old_path}/{dir}")
  cat(message)
  cat("\n")

  rmd_files <- append(rmd_files, glue("{old_path}/{dir}/{to_be_tested}"))
}
```

## List libraries

```{r}
#| label: list-libraries

lib <- NULL
for (i in 1:length(rmd_files)) {
  read_lines(rmd_files[i]) |>
    str_subset("library\\(") |>
    append(lib) -> lib
}
lib |>
  unique() |>
  sort() -> lib
cat(lib, sep="\n")
```

## Render files

```{r}
#| message: false
#| warning: false
#| label: render-files

for (i in 1:length(rmd_files)) {
  print(glue("{i}: {rmd_files[i]}"))
  tmp <- quarto_render(input=rmd_files[i], quiet=TRUE)
}
```